### 2023-07-15 Notes.tex
Notes

%(query-replace-regexp "\\\\sqsubseteq^{sh}_{\\([^}]*\\)}" "\\\\shiftpre[\\1]")
% --- LNCS ---'
\documentclass{llncs}

\usepackage{hyperref}
\usepackage[textsize=tiny]{todonotes}
\usepackage{etex}

\usepackage[all]{xy}
\SelectTips{cm}{}
% derivations and labels
\usepackage{proof}
\newcommand{\lab}[1]{\ensuremath{\mathsf{{#1}}}}
\newcommand{\slab}[1]{\ensuremath{\scriptstyle{\mathsf{{#1}}}}}

% ``boxed'' \infer command
\newcommand{\binfer}[3][]{
  \mbox{\infer[#1]{#2}{#3}}}

% Command for labels on the left side of the rule
%       \inferL{<name>}{<post>}{<pre>}
% generates:
%             <pre>
%      <name> ------
%             <post>
%
\newlength{\myheight}
\newcommand{\inferL}[3]
  {\settoheight{\myheight}{\mbox{${#2}$}}
   \raisebox{\myheight}{{#1}}
   \makebox[1mm]{}
   \mbox{\infer{#2}{#3}}
}

\usepackage{amssymb,graphicx,epsfig,color}
%\usepackage[scriptsize]{subfigure}
\usepackage{subcaption}
\usepackage{wrapfig}

% re-stating 
%\usepackage{thm-restate}

% showing labels
%\usepackage[inline]{showlabels}

\usepackage{pgf}
\usepackage{tikz}
\usepackage{tikz-cd}
\usetikzlibrary{arrows,shapes,snakes,automata,backgrounds,petri,fit,positioning}
\tikzstyle{node}=[circle, draw=black, minimum size=1mm]
\tikzstyle{trans}=[font=\scriptsize]
\tikzstyle{lab}=[font=\small]

\newcommand{\pgfBox}{
  \begin{pgfonlayer}{background} 
    \fill[blue!2,thick,draw=black!50,rounded corners,inner sep=3mm] ([xshift=-1.5pt,yshift=-1.5pt]current bounding box.south west) rectangle ([xshift=1.5pt,yshift=1.5pt]current bounding box.north east);
  \end{pgfonlayer}
}

\newcommand{\Rrel}[1]   {\stackrel{{#1}}{\Longrightarrow}}
\newcommand{\oa}{\overline a}
\newcommand{\ob}{\overline b}
\newcommand{\oc}{\overline c}
\newcommand{\od}{\overline d}
\newcommand{\nil}{{\tt 0}}
\newcommand{\rec}{\emph{rec}}
\newcommand{\fn}[1]{{\mathtt{fn}}(#1)}
%\usepackage{latexsym}
\usepackage{stmaryrd}
\def\encodep#1{\llfloor#1\rrfloor}

\newcommand{\cat}[1]{\ensuremath{\mathbf{#1}}}

\newcommand{\dpo}{\textsc{dpo}}

% base classes of categories for adhesive and quasi adhesive case
\newcommand{\bAdh}{\ensuremath{\mathbb{B}}}
\newcommand{\bQAdh}{\ensuremath{\mathbb{QB}}}

%from pawel
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
%2 righe tolte da Fabio
\usepackage{amssymb}
%\usepackage{amsthm}
\usepackage{enumerate}
\usepackage{xspace}
\usepackage{amsfonts}
\usepackage{mathrsfs}
\usepackage{cite}
\usepackage{float}
\usepackage{fancybox}

%%%%%%%% MATHEMATICAL NOTATION %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%symbol for natural numbers
\newcommand{\nat}{\ensuremath{\mathbb{N}}}

% finite subset
\newcommand{\sfin}{\ensuremath{\subseteq_{\mathit{fin}}}}

% flattening of a multiset
\newcommand{\flt}[1]{\ensuremath{[\![{#1}]\!]}}

% compact elements
\newcommand{\compact}[1]{\ensuremath{\mathop{\mathsf{K}({#1})}}}

% principal ideal
\newcommand{\principal}[1]{\ensuremath{\mathop{\downarrow\!{#1}}}}

% ideal completion
\newcommand{\ideal}[1]{\ensuremath{\mathsf{Idl}({#1})}}

% complete prime elements
\newcommand{\pr}[1]{\ensuremath{\mathop{\mathit{pr}({#1})}}}
\newcommand{\wpr}[1]{\ensuremath{\mathop{\mathit{wpr}({#1})}}}

% irreducible elements
\newcommand{\ir}[1]{\ensuremath{\mathop{\mathit{ir}({#1})}}}

% difference of irreducible elements
\newcommand{\diff}[2]{\ensuremath{\delta({#1},{#2})}}

% immediate precedence

% abbreviation for event structure
% \newcommand{\esabbr}{event structure}
\newcommand{\esabbr}{\textsc{es}}
\newcommand{\esnabbr}{\textsc{esnb}}
\newcommand{\esnmabbr}{\textsc{esn}}
\newcommand{\eseqabbr}{\textsc{epes}}

% predecessor of an irreducible
\newcommand{\pred}[1]{\ensuremath{\mathit{p}({#1})}}

% irreducible elements in es domains
\newcommand{\esir}[2]{\ensuremath{\langle{#1}, {#2}\rangle}}

% equivalence classes [of irreducibles]
\newcommand{\eqclass}[2][]{\ensuremath{[{#2}]_{\scriptscriptstyle {#1}}}}
% union of the equivalence classes of the elements in a set
\newcommand{\eqclasscup}[2]{\ensuremath{{#2}_{\scriptscriptstyle {#1}}}}

\newcommand{\eqclassir}[1]{\ensuremath{\eqclass[\leftrightarrow^*]{#1}}}

% quotient of set wrt a relation
\newcommand{\quotient}[2]{\ensuremath{{#1}_{\scriptscriptstyle {#2}}}}

% category of event structures 
\newcommand{\es}{\ensuremath{\mathsf{ES}}}
% category of stable event structures 
\newcommand{\ses}{\ensuremath{\mathsf{sES}}}
% category of prime event structures 
\newcommand{\pes}{\ensuremath{\mathsf{pES}}}
% category of prime event structures with equivalence
\newcommand{\epes}{\ensuremath{\mathsf{epES}}}

% category of connected event structures 
\newcommand{\ces}{\ensuremath{\mathsf{cES}}}

% category of weak prime algebraic domains domains 
\newcommand{\WDom}{\ensuremath{\mathsf{wDom}}}
% category of domains
\newcommand{\Dom}{\ensuremath{\mathsf{Dom}}}
% category of prime algebraic domains
\newcommand{\PDom}{\ensuremath{\mathsf{pDom}}}

%%%%% NON BINARY CONFLICT

% category of event structures 
\newcommand{\esn}{\ensuremath{\mathsf{ES_{nb}}}}
% category of stable event structures 
\newcommand{\sesn}{\ensuremath{\mathsf{sES_n}}}

% category of connected event structures 
\newcommand{\cesn}{\ensuremath{\mathsf{cES_{nb}}}}

% category of prime event structures 
\newcommand{\pesn}{\ensuremath{\mathsf{pES_n}}}

% category of fusion domains 
\newcommand{\WDomb}{\ensuremath{\mathsf{wDom_b}}}
% category of domains
\newcommand{\Domb}{\ensuremath{\mathsf{Dom_b}}}
% category of prime algebraic domains
\newcommand{\PDomb}{\ensuremath{\mathsf{pDom_b}}}

%%%%% END NON BINARY CONFLICT

% slice category
\newcommand{\slice}[2]{\ensuremath{({#1} \downarrow {#2})}}

% event structure for a domain
\newcommand{\zev}[0]{\ensuremath{\mathcal{E}}}
\newcommand{\ev}[1]{\ensuremath{\zev({#1})}}

% from general to connected event structures
\newcommand{\zconnes}[0]{\ensuremath{\mathcal{C}}}
\newcommand{\connes}[1]{\ensuremath{\zconnes({#1})}}
% and inclusion
\newcommand{\zinces}[0]{\ensuremath{\mathcal{I}}}
\newcommand{\inces}[1]{\ensuremath{\zinces({#1})}}

% stable version
\newcommand{\zsev}[0]{\ensuremath{\mathcal{E}_S}}
\newcommand{\sev}[1]{\ensuremath{\zsev({#1})}}

% with equivalence
\newcommand{\zeveq}[0]{\ensuremath{\mathcal{E}_{eq}}}
\newcommand{\eveq}[1]{\ensuremath{\zeveq({#1})}}

% es with equivalence to es and vice
\newcommand{\zfuse}[0]{\ensuremath{\mathcal{M}}}
\newcommand{\fuse}[1]{\ensuremath{\zfuse({#1})}}
\newcommand{\zunf}[0]{\ensuremath{\zunf}}
\newcommand{\unf}[1]{\ensuremath{\mathcal{U}({#1})}}

% Winskel/Droste version
\newcommand{\zevwd}[0]{\ensuremath{\mathcal{E}_{wd}}}
\newcommand{\evwd}[1]{\ensuremath{\zevwd({#1})}}

% configurations of an event structure
\newcommand{\conf}[1]{\ensuremath{\mathit{Conf}({#1})}}
% finite configurations
\newcommand{\conff}[1]{\ensuremath{\mathit{Conf_F}({#1})}}

% product of the sets of minimal enablinsg
\newcommand{\pmin}[1]{\ensuremath{U_{#1}}}

% connectectedness of minimal enablinsg
\newcommand{\conn}[1]{\ensuremath{\stackrel{#1}{\frown}}}

% domain for an event structure or graph grammar
\newcommand{\zdom}[0]{\ensuremath{\mathcal{D}}}
\newcommand{\dom}[1]{\ensuremath{\zdom({#1})}}

\newcommand{\zdomeq}[0]{\ensuremath{\mathcal{D}_{eq}}}
\newcommand{\domeq}[1]{\ensuremath{\zdomeq({#1})}}

% partial order for a graph grammar
\newcommand{\poset}[1]{\ensuremath{\mathcal{P}({#1})}}

% stable version
\newcommand{\pdom}[1]{\ensuremath{\mathcal{D}_S({#1})}}
\newcommand{\ppdom}[0]{\ensuremath{\mathcal{D}_S}}

% powerset
\newcommand{\Pow}[1]{\ensuremath{\mathbf{2}^{#1}}}

% powerset of finite subsets
\newcommand{\Powfin}[1]{\ensuremath{\mathbf{2}_\mathit{fin}^{#1}}}

% powerset of subsets of cardinality <= 1
\newcommand{\Powone}[1]{\ensuremath{\mathbf{2}_1^{#1}}}

% integer interval
\newcommand{\interval}[2][1]{\ensuremath{[{#1},{#2}]}}

% domain interval
\newcommand{\dint}[2]{\ensuremath{[{#1},{#2}]}}

% set of intervals
\newcommand{\IntSet}[1]{\ensuremath{\mathop{\mathit{Int}({#1})}}}

% intervals to irreducibles and vice
\newcommand{\inir}{\ensuremath{\mathop{\mathit{\zeta}}}}
\newcommand{\irin}{\ensuremath{\mathop{\mathit{\iota}}}}

% permutations
\newcommand{\perm}{\sigma}

% causes
\newcommand{\causes}[1]{\ensuremath{\lfloor {#1})}}

%%% GRAPH GRAMMARS

\newcommand{\Abs}[1]{\ensuremath{\mathsf{Abs}({#1})}}
\newcommand{\tr}[1]{\ensuremath{\mathsf{Tr}({#1})}}
% fusion safe traces
\newcommand{\trs}[1]{\ensuremath{\mathsf{Tr}_s({#1})}}
%\newcommand{\graph}{\ensuremath{\mathsf{Graph}}}
\newcommand{\tgraph}[1]{\ensuremath{\mathsf{Graph}_{#1}}}
\newcommand{\can}[1]{\ensuremath{\mathsf{C}({#1})}}
% source and target of a derivation
\newcommand{\source}[1]{\ensuremath{\mathsf{s}({#1})}}
\newcommand{\target}[1]{\ensuremath{\mathsf{t}({#1})}}
\newcommand{\col}[1]{\ensuremath{\mathsf{col}({#1})}}

% left decorated trace
\newcommand{\ltrace}[1]{\ensuremath{\langle {#1}\rangle_c}}

\newcommand{\bx}[1]{\phantom{\big(}#1{\phantom{\big)}}}
\newcommand{\bxx}[1]{\,#1\,}
\newcommand{\cycl}[1]{\ensuremath{\mbox{\textcircled{\scriptsize{$#1$}}}}}
\renewcommand{\iff}{\ensuremath{\Leftrightarrow}}

%%% NEW

\usepackage{xparse}

% inversions
\newcommand{\inv}[1]{\ensuremath{inv}({#1})}

% direct shift
\newcommand{\shiftdir}[1][]{\ensuremath{\mathrel{{\rightsquigarrow}^{\mathit{sh}}_{#1}}}}

% shift preorder
\newcommand{\shiftpre}[1][]{\ensuremath{\mathrel{{\sqsubseteq}^{\mathit{sh}}_{#1}}}}

% shift equivalence
\newcommand{\shifteq}[1][]{\ensuremath{\mathrel{{\equiv}^\mathit{sh}_{#1}}}}

% transp{source}[target]: if target not specified source+1
\NewDocumentCommand{\transp}{m o}{%
  \ensuremath{({#1},%
  \IfNoValueTF{#2}%
    {{#1}+1}%
    {#2}%
    )}
}

\NewDocumentCommand{\mycommand}{o}{%
  % <code>
  \IfNoValueTF{#1}
    {code when no optional argument is passed}
    {code when the optional argument #1 is present}%
  % <code>
}
% interchange
\newcommand{\IC}[1]{\ensuremath{\mathit{IC}({#1})}}

\title{Ideas}

\begin{document}
\maketitle

\subsection*{Shift equivalence for non-right linear}

Sequential independence is a key ingredient in the concurrency theory of DPO graph transformation, since it states when consecutive rewriting steps can be switched, thus 
identifying them as causally unrelated. The associated theory of shift equivalence on rewriting sequences is well-understood for linear rules, including its 
connection with standard models for concurrent systems such as event structures. 
%
This is not so for rules that are only left-linear, i.e., such that
their right-hand side may merge graph items.

In particular, if one just take the definition with general
interchange pairs, undesidered things might happen. In particular one
can have shift-equivalent derivations where the permutation is the
identity but which are non isomorphic (e.g., in GraphShift, page 1-2
$\rho_1 \shifteq \rho_3$, but $\rho_1 \equiv^{a} \rho_3$.

We want a notion of shift-equivalence which ensures some properties
which are indispensable to allow to interpret derivations up-to shift
as equivalent concurrent computations which differ only for the fact
that independent steps have been moved around. When moving around steps, we should not loose track of the identity of the steps: this is essential to be able to have an event structure semantics with a clearly defined notion of event.

Hence we assume to deal
with a notion of shift-equivalence between derivations of the kind
$\rho \shifteq_\sigma \rho'$ where $\rho, \rho'$ consist of $n$
rewriting steps, $\sigma: \interval{n} \to \interval{n}$ is a
permutation (the idea being that $\rho$ is obtained from $\rho'$
pemutating the steps according to $\sigma$) with the following
properties:

\begin{itemize}
\item \textbf{Property 0}. Shift equivalence is an equivalence, in particular it is symmetric. I.e., if  $\rho \shifteq_\sigma \rho'$ then $\rho' \shifteq_{\sigma^{-1}} \rho$;

\item \textbf{Property 1}. Derivations which are shift-equivalent via the
  identity permutation are isomorphic: if $\rho \shifteq \rho'$ then
  $\rho \equiv^a \rho'$;

\item \textbf{Property 2}. Shift equivalence between adjacent rewriting steps
  is a property of the steps, invariant for permutation according to
  shift, i.e., if two steps are independent in some derivation $\rho$
  and we can prove that $\rho$ is shift-equivalent to some $\rho'$
  where the same steps are again consecutive, but possibly in a
  different position, then the steps are again independent, and also
  the converse holds. Formally, if $\rho \shifteq_\sigma \rho'$ with
  $\rho, \rho'$ consist of $n$ rewriting steps and for some
  $i \in \interval{n-1}$, $\sigma(i)+1 = \sigma(i+1)$, the the steps
  $i, i+1$ in $\rho$ are shift-independent iff $\sigma(i)$ and
  $\sigma(i)+1$ are so in $\rho'$.
\end{itemize}

\textbf{Question.} Taking the standard shift, with generic interchange pair, a a counterexample to property 1, is the one mentioned above.  Do we have a counterexample to property 2? Not clear to me.

The above discussion should lead to

\begin{enumerate}
\item The conditions proposed by Andrea for a good notion of shift (in
  the draft, GraphShift, Def. 3, cond (1)-(5), with the addition of
  Lemma 12).

  \textbf{Questions}: Can we reduce the set of axioms? Do we need something else relating abstraction equivalence (iso of derivation) and shift? Relation with symmetries?

\item For left-linear rules, the proposed notion of shift equivalence
  requires: uniqueness of the interchange pair and symmetry. Show that
  this satisfies the above

\item Apparently, the satisfaction of the properties above has not
  been explicitly observed for the case of linear rules. If so, it is
  worth to remark it.

\item Does the theory above apply to other cases. E.g., if one wants a
  proper notion of causality (avoiding or-causality) a possibility
  seems to be to restrict the applicability condition for rules, to
  prevent fusions to act on already merged items (see section V.D in
  the arxiv). By the way, note that the satisfaction of the conditions
  on shift depends not only on the notion of shift, but also on the
  notion of rewriting. In this case we expect to have a prime event
  structure (bit this has to do with the second part of the work).

\end{enumerate}

\subsection*{Event structures for non-right linear}

Ideally, one would like to be able to work with an abstract notion of
shift-equivalence and just using its properties as singled out in the
previous part derive a weak prime domain or connected event structure
for it. This is probably difficult, but one might need to add some
futher properties to be proved for the specific notion of shift at
hand.

For this part, relevant material is section V of Arxiv. We use abstraction equivalence (iso of derivations) $\equiv^{a}$ (def. 36), shift-equivalence $\shifteq_\sigma$ (Def. 38, traking the permutation) and their join $\equiv^c_\sigma$.

For linear rules on graphs it is known that two derivations are
related by $\rho \equiv^c_\sigma \rho'$ iff $\sigma$ is a consistent
permutation (def. 40, arxiv). Rough;y the idea is to take the colimit
and consider the rewriting steps over the colimit to deduce
dependencies. This was used in many places in the proofs, but for
left-linear rules the proof does not go through (only the implication
``if $\rho \equiv^c_\sigma \rho'$ then $\sigma$ is a consistent
permutation'' (Lemma 17(1)) seems to hold as the original proof do not
to rely on linearity, this should be double-checked, but the converse
is false). In general, the hope is to get rid of using consistent
permutation and use instead more asbtract properties of
shift-equivalence as discussed in the first part.

The general idea is to show that considering derivations are arrows from the source to the target graph, then the slice from the initial graph $\slice{[G_s]}{\tr{\mathcal{G}}}$ should be a pre-order, whose ideal completion is a weak prime domain.

Relevant technical lemma in Arxiv are:

\begin{itemize}

\item
  \textbf{Lemma 17}. Points (2) and (3) should be hopefully consequences of the general properties of shift in the previous part and they play a role in proving that $\slice{[G_s]}{\tr{\mathcal{G}}}$ is a pre-order (Prop. 14)

\item
  \textbf{Lemma 19}.
  (1) says that if $\psi; \psi_1 \equiv^c_\sigma \psi'; \psi_1'$ and the permutation $\sigma$ sends $n$ steps of $\psi_1$ to $\psi'$ then such steps are not in $\psi$ and they will need to be in every $\phi_2$ such that $\psi; \phi_2 \equiv^c_\sigma \psi'; \phi_2'$. In particular there is $\phi_2$ consisting of exactly $n$ steps which allows for commutation. This allows essentially to deduce existence of minimal upper bounds, al with the same length. 
  Then (2) shows that all minimal upper bounds are isomorphic, so that we have a sup.

  Again, the hope is to obtain this from the abstract properties of
  shift.

\item \textbf{Proposition 15} shows that $\ideal{\poset{\mathcal{G}}}$
  is a domain, i.e., algebraic finitary (see Def. 6), while
  \textbf{Theorem 6} prove that it is weak prime.

  Here I think that the proofs are draft modifications of the original ones, changed after discovering the property on consistent permutation to be wrong. But they are drafty, to be redone maybe keeping some ideas.

\subsection*{Event structures for non-right linear, but jointly mono matches}

This is the idea in Section V.D. As mentioned above, ideally, if the proof in the previous part were based just on abstract properties of shift, one could hope of proving that shift in this context has the right properties, deduce that the construction produces a weak prime domain.

Then, one could observe that in this specific context actually all weak primes are primes an thus deduce that this provides a prime algebraic domain/pes semantics of fusion.

\end{itemize}

\end{document}

\keywords{Shift equivalence, non-linear rules, DPO rewriting.}

\section{Introduction}
\label{se:graphs}

Shortly, shift equivalence can be defined as the smallest equivalence relation on derivations, i.e. sequences of rewriting steps,
that can be obtained by the iterated switch of sequentially 
independent steps. A crucial observation, which represented our starting point, is that the notion of sequential independence traditionally
adopted for linear graph transfomation systems is not adequate when we
allow rules to be non-right linear, i.e., in the presence of
fusions. Consider the rules in Fig.~\ref{fi:rules}, typed over the type graph $T$,
where numbers are used to represent the mappings from the interface to the left- and
right-hand sides.

\begin{figure}
  \input{Figure/rules.tex}

\caption{The type graph  and the rules of the grammar of the example.}
\label{fi:rules}
\end{figure}

Now, let us consider an initial graph with a single node. We can then obtain
the derivation $\rho_1 = \delta_1; \delta_2; \delta_3$ in Fig.~\ref{fi:der1}, where $\delta_2$ and $\delta_3$ are
sequential independent according to the standard notion.

\begin{figure}
  \input{Figure/der1.tex}
\caption{The derivation $\rho_1$.}
\label{fi:der1}
\end{figure}

We can exchange the applications of $p_2$ and $p_3$, thus getting the
derivation $\rho_2 = \delta_1; \delta_3'; \delta_2'$ in
Figure~\ref{fi:der2}. Now, one can exchange again $p_3$ and $p_2$,
reaching the derivation $\rho_3 = \delta_1; \delta_2'; \delta_3$ in
Figure~\ref{fi:der3}, where in $\delta_2'$ production $p_2$ uses the
node of the start graph and thus ``does not depend'' on $p_1$.

\begin{figure}[t]
  \input{Figure/der2.tex}
\caption{The derivation $\rho_2$.}
\label{fi:der2}
\end{figure}

\begin{figure}[t]
  \input{Figure/der3.tex}
\caption{The derivation $\rho_3$.}
\label{fi:der3}
\end{figure}

The fact that if we switch two steps and then switch them back, we obtain a derivation that is not equivalent to the original one has unpleasant consequences. For instance, if we restrict $\rho_1'$ and $\rho_3'$ to the first two steps, we get non-equivalent derivations $\delta_1;\delta_2 \not\shifteq \delta_1;\delta_2'$, but $\rho_1 \shifteq[\sigma] \rho_3$, with $\sigma$ mapping $\delta_1$ to $\delta_1$ and $\delta_2$ to $\delta_2'$.
%
The lack of such closure with respect to composition does not allow to obtain a causal semantics for graph transformation systems.
More precisely, it does  not allow to associate an event structure to a derivation.
%
Thus, the notion of sequential independence has to be refined. The point is that in the associated shift equivalence we want to forget who made the fusion, 
but we should not forget (as it happens in the example) that some fusions occurred.

% In Andrea's counterexample (same as above, but $p_2$ produces an item that is consumed by $p_3$, thus forcing $p_3$ to happen last), if we construct the domain we have:
% \begin{center}
%   \includegraphics[scale=.7]{dom.pdf}
% \end{center}
% which is a weak prime domain (actually a prime domain) but it does not capture the proper semantics (e.g., $p_2'$ and $p_3$ are the same event!).

Shift equivalence is made stricter by asking that two steps are
sequentially independent if the result of their switch is uniquely
determined. Intuitively, when there are several way of performing the
exchange, it means that the first step has executed some fusions that
the second step is using in an essential way, hence they should not
considered independent. Formally, this boils down to require that 
in the definition of sequential independence the interchange pair is unique.

The paper builds on this intuition: it introduces this stricter notion of sequential independence
(Section~xxx) and prove that the associated theory of shift equivalence on derivations is 
well-behaved with respect to the concatenation of rewriting steps (Section~xxx).
Finally, it outlines how the relationship with event structure works, wrapping up the paper
with comparison with related woks and hinting at further direction of research (Section~xxx).

\section{Preliminaries [to be redone]}

\begin{figure}[t]
   \input{Figure/basic_independence}
\caption{Sequential independence for
  $\rho = G \Rrel{p_1/m_1} H \Rrel{p_2/m_2} M$.}
\label{fi:strongseq}
\end{figure}

\begin{definition}[sequential independence]
  \label{de:seq-ind}
  Consider a derivation $G \Rrel{p_1/m_1} H \Rrel{p_2/m_2} M$ as in
  Fig.~\ref{fi:strongseq}. Then, its components are \emph{sequentially
    independent} if there exists a unique \emph{independence pair} among
  them, i.e., two graph morphisms $i_1: R_1 \rightarrow D_2$ and
  $i_2: L_2 \rightarrow D_1$ such that $l_2^* \circ i_1 = m_{R_1}$,
  $r_1^* \circ i_2 = m_{L_2}$.
\end{definition}

Observe that this notion reduces to the usual one for linear rules. In
fact when $r_1^*$ and $l_2^*$ are mono the uniqueness requirement is
trivially satisfied. Instead, for left-injective rules $i_1$ is still
unique, but $i_2$ might be not (as it happens, e.g., in the example
before).\todo{Say more about the example}

Observe that the notion is not symmetric: if
$\rho = G \Rrel{p_1/m_1} H \Rrel{p_2/m_2} M$ is sequential independent
and we switch it, we could obtain a pair which is not because it fails
to satisfy uniqueness of the independence pair (this happens, e.g., in
our example, where $\delta_2;\delta_3$ is sequential independent, but
$\delta_3';\delta_2'$ is not). The idea is to define shift equivalence
as $\rho \equiv_{sh} \rho'$ when \emph{each of the two} can be
obtained from the other by exchanging sequential independent
derivations. More precisely, we have a shift relation, which is used
for defining the equivalence as in Fig.~\ref{fi:ind}. It could be
better to have a slightly less elegant but more direct
definition. Write $\rho \shiftdir[\transp{i}] \rho'$ when
$\rho = \rho_1; \rho_2; \rho_3$ and $\rho' = \rho_1; \rho_2', \rho_3$
with $|\rho_1| = i-1$, $|\rho_2|=2$ and $\rho_2' \in
\IC{\rho_2}$. Then we write $\rho \shiftpre[\perm] \rho'$ if
$\rho = \rho_0 \shiftdir[\perm_1] \rho_1 \ldots \shiftdir[\perm_n ]
\rho'$ with $\perm = \perm_1; \ldots; \perm_n$ for some $n \geq 0$
(with the convention that for $n=0$ we get the identity).

The restricted notion of sequential independence ensures that the interchange operator produces a result that is unique up to abstraction equivalence.

\begin{lemma}[IC is unique]
  \label{le:unique-IC} 
  Let $\rho$ be a derivation such that $|\rho|=2$. If $\rho$ is
  sequential independent, let us write
  $\IC{\rho} = \{ \rho' \mid \rho \shiftdir \rho'\}$. Then for
  all $\rho', \rho'' \in \IC{\rho}$ we have $\rho' \equiv
  \rho''$. Moreover, if $\rho$ is sequential independent then for all
  $\rho' \in \IC{\rho}$, if $\rho'$ is sequential independent and
  $\rho'' \in \IC{\rho'}$, we have $\rho \equiv \rho''$.
\end{lemma}

\begin{figure}[t]
  \input{Figure/independence}
  \caption{Independence.}
  \label{fi:independence}
\end{figure}

It is convenient (at the beginning, at least) to work with shift
on derivations where the source and target graph is concrete and the
middle can be taken up to iso. I think that everything can be then
easily transferred to ctc-equivalence. We say that $\rho$ and $\rho'$
are abstract shift-equivalent and write $\rho \sim \rho'$.

A crucial observation is that independence can be defined also for non-contiguous steps.

\begin{definition}[independence]
  Let $\delta_1;\rho;\delta_n$ be a derivation as in
  Fig.~\ref{fi:independence}, where $\delta_1$ and $\delta_n$ are
  direct derivations and $\rho$ is a generic, possibly empty
  derivation. We say that $\delta_1$ and $\delta_n$ are independent if
  there is a unique independence pair $i_{n,1}$ and $i_{1,n}$
  satisfying the expected commutation, where $C$ is the colimit of the
  bottom line (excluding $G_1, G_{n+1}$).
\end{definition}

The fact that the notion of independence makes sense is proved by the following lemma, showing that the notion is preserved by the shift relation.

\begin{lemma}[independence is global]
  \label{le:seq-ind-glob1}
  The notion of independence for contiguous derivation steps
  coincides with sequential independence.

  Moreover, independence is invariant under the shift-relation, i.e.,
  \begin{itemize}

  \item if $\delta_1;\rho;\delta_2$ and $\rho \shiftpre \delta'$ then
    in $\delta_1;\rho;\delta_2$ the steps $\delta_1$, $\delta_2$ are
    independent iff they are in $\delta_1;\delta';\delta_2$.

  \item if $\delta_1;\rho;\delta';\delta_2$ and
    $\delta_2'; \delta'' \in \IC{\delta';\delta_2}$ then in
    $\delta_1;\rho;\delta';\delta_2$ the steps $\delta_1$ and $\delta_2$ are
    independent iff they are in $\delta_1;\rho;\delta_2'$.

  \item same as before, but with shift on the left.
  \end{itemize}
\end{lemma}

\begin{proof}
  To be done. The fact that sequential independence coincides with
  independence for contiguous derivations is easy. The fact that it coincides when derivations are not contiguous is false (the approach based on defining independence for non-contiguous derivation is not working, but the results should be true ...)
\end{proof}

A \emph{permutation} on $\interval{n}$ is a bijection
$\perm : \interval{n} \to \interval{n}$. We call it a
\emph{transposition} when there are $i, j \in \interval{n}$,
$i \neq j$ such that $\sigma(i)=j$, $\sigma(j) = i$ and
$\sigma(k) = k$ otherwise. Such a transposition is normally denoted as
$\transp{i}[j]$. Whenever $j = i+1$ we call it an \emph{adjacent
  transposition}.

Given a permutation $\perm$ on $\interval{n}$, the set of inversions
of $\perm$ is
$\inv{\perm} = \{ (i,j) \mid i,j \in \interval{n-1}\ \land\ i<j\
\land\ \perm(i) > \perm(j) \}$. Clearly, $\perm$ is the identity iff
$\inv{\perm}=0$.

Now, if $\perm \neq id$ there is at least an inversion of the kind
$(i,i+1)$.  Then we have that $\perm = \transp{i}; \perm'$
with $|\inv{\perm'}| = |\inv{\perm}| -1$. Hence we can inductively
express $\perm$ as a (minimal length) composition of adjacent
transpositions.

\begin{lemma}[on permutations]
  \label{le:perm}
  Let $\perm$, $\perm'$ be permutations on $\interval{n}$ and
  $\perm = \upsilon_1; \perm'; \upsilon_2$ where $\upsilon_1$ and $\upsilon_2$
  are transpositions with $\upsilon_1 = \transp{i}[j]$ and
  $\upsilon_2 = \transp{\perm'(i)}[\perm'(j)]$. Then $\sigma =
  \sigma'$.

  Moreover, let $\upsilon_1, \upsilon_2$ be adjacent
  transpositions and $\sigma' = \upsilon_1; \ldots, \upsilon_m$ a
  minimal length decomposition of $\sigma'$ in terms of adjacent
  transpositions (that thus operate on inversions). Then, for all
  $h \in \interval{m}$ with $\upsilon_h = \transp{l}[l+1]$ there exists
  $k \in \interval{m}$ such that
  $\upsilon_k = \transp{\upsilon_{1,h-1}^{-1};v_1;
    \upsilon_{1,k-1}(l)}[\upsilon_{1,h-1}^{-1};v_1; \upsilon_{1,k-1}(l+1)]$.
\end{lemma}

\begin{proof}
  The first part is easy. The second is a consequence of the fact that
  the minimal decomposition in terms of adjacent transpositions is
  essentially unique, in the sense that it consists of exchanges of
  all the inversions. To be formalised.
\end{proof}

In words, the lemma above says that if a permutation is decomposed as $\sigma = \upsilon_1; \sigma'; \upsilon_2$ with $\sigma' = \upsilon_1'; \ldots, \upsilon_n'$, where $\upsilon_1$ and $\upsilon_2$ performs a switch and its reverse, and the inner decomposition is minimal, then we can strip the first and last switch and get a decomposition of the same permutation that performs the same switches but possibly in a different order.

\begin{lemma}[no need of useless shifts]
  \label{le:no-useless-shifts}
  Let $\rho$, $\rho'$ be direct derivations and assume
  $\rho \shiftpre[\sigma] \rho'$. Then
  $\rho = \rho_0 \shiftdir[\upsilon_1] \rho_1 \ldots \shiftdir[\upsilon_n ]
  \rho'$ where $\sigma = \upsilon_1; \ldots; \upsilon_n$ and $\upsilon_j \in \inv{\upsilon_{j,n}}$ for $j \in \{1, \ldots,n\}$.
\end{lemma}

\begin{proof}
  Let $\rho \shiftpre[\perm] \rho'$. This means that
  $\rho = \rho_0 \shiftdir[\upsilon_1] \rho_1 \ldots \shiftdir[\upsilon_n ]
  \rho'$ with $\perm = \upsilon_1; \ldots; \upsilon_n$ for some $n \geq 0$.

  We show that if some $\upsilon_j$ does not act on an inversion, i.e.,
  $\upsilon_j \not\in \inv{\sigma_{j,n}}$ then we can shorten the
  proof of $\rho \shiftpre[\sigma] \rho'$ of two steps. This will
  allow us to conclude that we can reduce it to a proof only acting
  on inversions, as desired.

  The key observation is that, if
  $\upsilon_j \not\in \inv{\sigma_{j,n}}$ then, intuitively, the ``same
  exchange'' must be performed in the reverse direction by some
  $\upsilon_h$ ($h>j$).  Formally, if $\upsilon_j = \transp{i}$ and
  $(i,i+1)\not\in\inv{\upsilon_{i,n}}$ then there exists $h>j$ such that
  $\upsilon_{j,h-1}(i) = \upsilon_{j,h-1}(i+1)+1$ and
  $\upsilon_h = \transp{\upsilon_{j,h-1}(i+1)}$. Among all the pairs
  $j, h$ of this kind, take one with minimal $h-i$. This means that in
  the middle all transpositions are inversions of $\upsilon_{i+1,h-1}$.

  By Lemma~\ref{le:perm}, $\upsilon_i$ and $\upsilon_h$ can be
  stripped, i.e., $\upsilon_{i,h} = \upsilon_{i+1,h-1}$. Moreover,
  again by Lemma~\ref{le:perm},
  $\upsilon_{i+1}, \ldots, \upsilon_{h-1}$ perform exchanges that were
  already done in the original proof, and thus, since independence is
  invariant under shift equivalence, as expressed by
  Lemma~\ref{le:seq-ind-glob1} they are allowed. Moreover, since by
  Lemma~\ref{le:unique-IC} the result of switching is uniquely
  determined (up to iso), the resulting derivation is the same as
  before, hence we can strip $\upsilon_{i}$ and $\upsilon_{h}$,
  and conclude.
\end{proof}

Then we can prove that whenever $\rho \shiftpre[\perm] \rho'$, each
decomposition of $\perm$ in terms of inversions leads to a valid
derivation of $\rho'$ from $\rho$.

\begin{lemma}[independence is global/2]
  \label{le:seq-ind-glob2}
  Let $\rho$, $\rho'$ be direct derivations. Let
  $\rho \shiftpre[\perm] \rho'$, with $\perm \neq id$ and let
  $\perm = \transp{i}; \perm'$ where $(i,i+1) \in \inv{\perm}$. Then
  there exists $\rho''$ such that
  $\rho \shiftpre[\transp{i}] \rho'' \shiftpre[\perm']
  \rho'$.
\end{lemma}

\begin{proof}
  Let
  $\rho = \rho_0 \shiftdir[\upsilon_1] \rho_1 \ldots
  \shiftdir[\upsilon_n] \rho'$ where $\sigma = \upsilon_1; \ldots; \upsilon_n$ and $\upsilon_j \in \inv{\upsilon_{j,n}}$ for $j \in \{1, \ldots,n\}$, be the decomposition
  provided by Lemma~\ref{le:no-useless-shifts}.

  We know that steps corresponding to $i$ and $i+1$ are exchanged at
  some point, namely there is $j$ such that
  $\upsilon_j =
  \transp{\upsilon_{1,j-1}(i)}[\upsilon_{1,j-1}(i+1)]$. By
  Lemma~\ref{le:seq-ind-glob1} the steps $i$ and $i+1$ are independent
  already at the beginning and thus we can exchange them first.
  More precisely, 
  \begin{center}
    $\rho \shiftdir_{\transp{i}} \rho'' \shiftdir_{\upsilon_{1,j-1}}
    \rho_j \shiftdir_{\upsilon_{j+1,n}} \rho'$.
  \end{center}
  This can be proved by induction on $j$.
\end{proof}

\begin{lemma}[more on shift]
  \label{le:more-shift}
  if $\rho_1;\rho_1' \shiftpre[\perm | \perm'] \rho_2; \rho_2'$,
  with $|\rho_1| = |\rho_2|$ and
  $\sigma : \interval{|\rho_1|-1]} \to \interval{|\rho_2|-1]}$ then
  $\rho_1 \shiftpre[\perm] \rho_2$ and
  $\rho_1' \shiftpre[\perm'] \rho_2'$.
\end{lemma}

\begin{proof}
  The results follows by the fact that, from
  Lemma~\ref{le:no-useless-shifts}, we derive that
  \begin{center} 
    $\rho_1;\rho_1' \shiftdir[ \upsilon_1 \mid id_{[1..|\rho_1'|]}]
    \ldots \shiftdir[ \upsilon_h \mid id_{[1..|\rho_1'|]}] \rho_2;
    \rho_1' \shiftdir[ id_{[1..|\rho_1|]} \mid \upsilon_1'] \ldots
    \shiftdir[ id_{[1..|\rho_1|]} \mid \upsilon_k'] \rho_2; \rho_2'$
  \end{center}
  where $\upsilon_1; \ldots; \upsilon_h$ is a decomposition of $\perm$
  in terms of inversions and, similarly,
  $\upsilon_1'; \ldots; \upsilon_k'$ is a decomposition of
  $\perm'$. Then
  $\rho_1 \shiftdir[\upsilon_1] \ldots \shiftdir[\upsilon_h] \rho_2$
  and analogously
  $\rho_1' \shiftdir[\upsilon_1'] \ldots \shiftdir[\upsilon_k']
  \rho_2'$, which leads to the desired conclusion.
\end{proof}

\todo[inline]{Note that above we assume that $\rho_1$ and $\rho_1'$ ends in the same graph. To be formalised.}

\begin{lemma}[uniqueness of permutation]
  if $\rho \shiftpre[\perm_1] \rho'$ and
  $\rho \shiftpre[\perm_2] \rho'$ then $\perm_1 = \perm_2$. In
  particular, if $\rho_1 \shiftpre[\perm_1] \rho_1$ then $\perm_1 = id$.
\end{lemma}

\begin{proof}
  We proceed by induction on the length of the derivations
  $\ell=|\rho|=|\rho'|$ and by an inner induction on the number of
  inversions in $\sigma_1$. When $\ell = 0$, the result is
  trivial. Let us assume $\ell>0$ and let $i_1 = \sigma_1^{-1}(1)$,
  $i_2 = \sigma_2^{-1}(1)$. Let
  $L_1' \stackrel{l_1'}{\leftarrow} K_1' \stackrel{r_1'}{\rightarrow}
  R_1')$ be the rule used by the first production in $\rho'$.
  By Lemma 17(1) in the paper [the implication
  that holds] $\sigma_1$ and $\sigma_2$ are consistent permutations of
  $\rho$ into $\rho'$.
  %
  Since they are consuming, it is easy to see that
  $i_1=\sigma_1^{-1}(1)$ is uniquely determined by the (equivalence
  class) of an item in the initial graph consumed by $L_1'$ and the
  same applies to $i_2=\sigma_2^{-1}(1)$, hence we conclude that
  $i_1=i_2$.

  Now, if $i_1=i_2=1$, then $\sigma_j = \transp{1}[1] \mid \sigma_j'$
  for $j \in \{1,2\}$ and $\rho = \delta; \rho_1$,
  $\rho' = \delta; \rho_1'$ where $\delta$ is the first direct
  derivation, and $\rho_1 \shiftpre[\perm_j'] \rho_1'$ for
  $j \in \{1,2\}$. Since $|\rho_1| = \ell-1$ we conclude by inductive
  hypothesis that $\sigma_1'=\sigma_2'$ and thus $\sigma_1=\sigma_2$,
  as desired.

  If instead $i_1=i_2=i>1$, since $\transp{i_1-1}[i_1]$ is an inversion for both
   $\sigma_1$ and $\sigma_2$, by Lemma~\ref{le:seq-ind-glob2} we can
  perform such a shift first, obtaining
  \begin{center}
    $\rho \shiftdir[{\transp{i-1}[i]}] \rho_1
    \shiftpre[{\transp{i-1}[i];\sigma_1}] \rho'$
    %
    \quad and \quad
    %
    $\rho \shiftdir[{\transp{i-1}[i]}] \rho_1
    \shiftpre[{\transp{i-1}[i];\sigma_2}] \rho'$
  \end{center}
  Since ${\transp{i-1}[i];\sigma_1}$ has an inversion less than
  $\sigma_1$ we conclude by inductive hypothesis that
  ${\transp{i-1}[i];\sigma_1} ={\transp{i-1}[i];\sigma_2}$, whence
  $\sigma_1=\sigma_2$.
\end{proof}

Note that the above works also with the liberal definition of
sequential independence.

\begin{lemma}[Lemma 17(2) in the paper]
  if $\rho;\rho_1 \shiftpre[\perm] \rho; \rho_2$ then
  $\perm = id_{[1..|\rho|]} \mid \perm'$ and
  $\rho_1 \shiftpre[\perm'] \rho_2$.
\end{lemma}

\begin{proof}[skecth]
  Productions are `uniquely identified by the items they consume,
  hence productions in $\rho$ are necessarily mapped to themselves,
  i.e., $\perm = id_{[1..|\rho|]} \mid \perm'$. Then we conclude
  $\rho_1 \shiftpre[\perm'] \rho_2$ by Lemma~\ref{le:more-shift}.
\end{proof}

\begin{lemma}[Lemma 20 in the paper]
  \label{le:prop-c}
  Let $\psi$ and $\psi'$ be decorated derivations. Then the following hold:
  \begin{enumerate}
  \item
    \label{le:prop-c:1}
    Let $\psi_1, \psi_1'$ be such
    that $\psi;\psi_1 \equiv^c_{\perm} \psi';\psi_1'$ and let
    $n = | \{ j \in \interval[|\psi|]{|\psi;\psi_1|-1} \mid \perm(j)
    < |\psi'| \}|$. Then for all $\phi_2, \phi_2'$ such that
    $\psi;\phi_2 \equiv^c \psi';\phi_2'$ it holds $|\phi_2| \geq n$
    and there are $\psi_2, \psi_2',\psi_3$ such that

    \begin{itemize}
    \item $\psi;\psi_1 \equiv^c \psi;\psi_2;\psi_3$
    \item $\psi;\psi_2 \equiv^c \psi';\psi_2'$

    \item $|\psi_2| = n$

    \end{itemize}

  \item
    \label{le:prop-c:2}
    Let $\psi_1, \psi_1'$, $\psi_2, \psi_2'$ be such that
    $\psi;\psi_1 \equiv^c_{\perm_1} \psi';\psi_1'$ and
    $\psi;\psi_2 \equiv^c_{\perm_2} \psi';\psi_2'$ with $\psi_1, \psi_2$
    of minimal length. Then $\psi_1 \equiv^c_\perm \psi_2 \cdot \nu$,
    where $\nu : \target{\psi_2} \to \target{\psi_2}$ is some graph isomorphism and
    $\perm(j) = \perm_2^{-1}(\perm_1(j + |\psi|))-|\psi|$ for
    $j \in \interval[0]{|\psi_1|-1}$.    
  \end{enumerate}
\end{lemma}

\begin{proof}[sketch]  
  Point ~\ref{le:prop-c:1}: The proof should be similar to the one we have already, basing the rearrangement on Lemma~\ref{le:seq-ind-glob2}. Roughly we get $\psi;\psi_1 \equiv^c \psi_a;\psi_b;\psi_{2};\psi_{3}$ and
  $\psi';\psi_1' \equiv^c \psi_a';\psi_b';\psi_{2}';\psi_{3}'$
  with
  \begin{center}
    $\psi_a;\psi_b;\psi_{2};\psi_{3}
    \equiv^c_{\perm_a|\perm_{b2}|\perm_2}
    \psi_a';\psi_b';\psi_{2}';\psi_{3}'$
  \end{center}
  with $\perm_a$ over $\interval{|\psi_a|-1}$, $\perm_{b2}$ over
  $\interval{|\psi_b;\psi_{2}|-1}$ ``mapping $\psi_b$ to $\psi_{2}'$
  and $\psi_b'$ to $\psi_{2}$'', $\perm_3$ over
  $\interval{|\psi_2|-1}$.  We deduce by Lemma~\ref{le:more-shift}
  that $\psi_a \equiv^c_{\perm_a} \psi_a'$,
  $\psi_b;\psi_{2} \equiv^c_{\perm_{b2}} \psi_b';\psi_{2}'$ and
  $\psi_{3} \equiv^c_{\perm_3} \psi_{3}'$.

  Now, for all $\phi_2, \phi_2'$ such that
  $\psi;\phi_2 \equiv^c \psi';\phi_2'$ we can perform the same
  splitting and this leads to the conclusion that $|\phi| \geq n$ and
  also it can be expressed as $\psi_2;\phi_2'$, that should subsume
  also the second point.
  %
\qed
\end{proof}

% \paragraph{Note.} This does not solve the problem raised by Andrea's counterexample that still applies.

Another idea could be to try to use the inductive characterisation in Figure~\ref{fi:ind} (but I do not see how to mange the transitive rule).

\begin{figure}
  \begin{center}
    \small
  \[
    \inferL{\slab{(SH-id)}}
    {\rho \shiftpre[id] \rho}
    {}
    % 
    \hspace{0.9cm}
    % 
    \inferL{\slab{(SH-exch)}}
    {\rho \shiftpre[{\transp{1}[2]}] \rho'}
    {\rho = \upsilon_1;\upsilon_2 & \rho' \in \IC{\rho}}
  \]

  \[
    \inferL{\slab{(SH-comp)}}
    {\rho_1 ; \rho_2  \shiftpre[\upsilon_1 \mid \upsilon_2] \rho'_1 ;
      \rho'_2}
    {\rho_1 \shiftpre[\upsilon_1] \rho'_1, \ \rho_2 \shiftpre[\upsilon_2] 
      \rho'_2,\ \tau(\rho_1) = \perm(\rho_2)}
  \]

  \[
    % 
    % \inferL{\slab{(SH-sym)}}
    % {\rho' \shiftpre[\perm^{-1}] \rho}
    % {\rho \shiftpre[\perm] \rho'}
    % % 
    % \hspace{0.3cm}
    % 
    \inferL{\slab{(SH-tr)}}
    {\rho \shiftpre[\perm'\circ \perm] \rho''}
    {\rho \shiftpre[\perm] \rho',\ \rho' \shiftpre[\perm'], \rho''}
    %
    \hspace{0.3cm}
    \inferL{\slab{(SH-eq)}}
    {\rho' \shifteq[\perm] \rho}
    {\rho \shiftpre[\perm] \rho' & \rho' \shiftpre[\perm^{-1}] \rho}
  \]

%  \noindent
%  for $\perm_1 \mid \perm_2$ the pairing of two permutations,
%  and similarly $\sqsubseteq^c_{\perm}$ is defined inductively by

  \[
    \inferL{\slab{(C-abs)}}
    {\psi \equiv^c_{id} \psi'} 
    {\psi \equiv^{a} \psi'} 
    % 
    \hspace{0.3cm}
    % 
    \inferL{\slab{(C-sh)}}
    {\langle \alpha, \rho, \omega \rangle \equiv^c_{\perm} \langle \alpha, \rho', \omega \rangle} 
    {\rho \shifteq[\perm] \rho'}
    % 
    \hspace{0.3cm}
    % 
    \inferL{\slab{(C-tr)}}
    {\psi \equiv^c_{\perm' \circ \perm} \psi''}
    {
      {\psi \equiv^c_{\perm} \psi'} & 
      {\psi' \equiv^c_{\perm'} \psi''}
    }
  \]
\end{center}
\caption{Inductive definition of $\shiftpre$ and $\equiv^c$, for $\perm_1 \mid \perm_2$ the pairing of two permutations.}
\label{fi:ind}
\end{figure}

% \paragraph{Consistent $5$-tuple}. The alternative is to try to amend the notion of consistent $5$-tuple by adding some dependencies between productions and asking that they are ``preserved'' by the mapping. With the new definition I fear that it could be difficult distinguishing between dependencies arising from fusions and other, hence one should try to capture all dependencies, that will concretise in a deterministic event structure. Concretely, given $\rho = p_1 p_2 \ldots p_n$ it could be mapped to $E(\rho) = \langle \{ 1, \ldots, n\}, \vdash\rangle$ where $X \vdash i$ if there exists $\rho \equiv_\perm \rho'$ such that for all $j \in X$ it holds $j < \perm(i)$.

\section{Proving invariance of independence with respect to shift}
\newcommand{\si}{\ensuremath{\mathrel{\langle \mbox{\textsc{si}}\rangle}}}

This is the way I (AC) propose to organize the proof that sequential independence is invariant with respect to shift. I copy and update some results of the previous part. Not everything is formalized completely.

The structure of the proof should be applicable to any notion of \emph{sequential independence} for which we can prove the following facts:

\begin{definition}[conditions on sequential independence]
\label{def:si-conditions}
\begin{enumerate}
\item \label{cond:CR} Sequential independent steps can be switched. If two consecutive steps $\delta_i$ and $\delta_{i+1}$ of a derivation $\rho$ are SI (sequential independent, written $\delta_i \si \delta_{i+1}$), then there is a constructive way to obtain a derivation $\rho'$ differing from $\rho$ only for steps $\delta'_i$ and $\delta'_{i+1}$, such that $rule(\delta_i) = rule(\delta'_{i+1})$ and $rule(\delta_{i+1}) = rule(\delta'_{i})$. In this case we write $\rho \shiftdir[i] \rho'$.

\item  \label{cond:CRdet} Switching is deterministic up to iso. If $\delta_i \si \delta_{i+1}$ in $\rho$ and $\rho \shiftdir[i] \rho'$, $\rho \shiftdir[i] \rho''$, then $\rho' \equiv \rho''$.

\item \label{cond:si-symm} Sequential independence is preserved by immediate switch. If $\delta_i \si \delta_{i+1}$ in $\rho$ and $\rho \shiftdir[i] \rho'$, then $\delta'_i \si \delta'_{i+1}$ in $\rho'$.

\item  \label{cond:si-si-iso} Switch is reversible. If $\rho \shiftdir[i] \rho'$ then $\rho' \shiftdir[i] \rho$.

\item As a consequence of the previous points, if $\rho \shiftdir[i] \rho'$ and $\rho' \shiftdir[i] \rho''$ then $\rho \equiv \rho''$.

\end{enumerate}
\end{definition}
%Given a derivation $\rho = \delta_1; \ldots ; \delta_i ; \delta_{i+1}; \ldots; \delta_n$, if $\delta_i \si \delta_{i+1}$, then
\todo[inline]{I think that all conditions are satisfied by the standard definition of SI in DPO with linear rules and by our stronger notion of SI (uniqueness of the two mediating morphism \textbf{and} same property of the switched steps) in DPO with possibly non-right-linear rules. }

\begin{proposition}[$\si$ satisfies all the conditions...]
Let's call the relation of Def.~\ref{de:seq-ind} \emph{weak sequential independence}. Then it satisfies conditions \ref{cond:CR} and \ref{cond:CRdet} of Definition~\ref{def:si-conditions} (by the classical Church-Rosser Theorem and uniqueness of the mediating morphisms), but not the remaining ones.

Let \emph{sequential independence} be defined as follows: $\delta_i \si \delta_{i+1}$ in $\rho$ iff they are weak sequential independent, and if $\rho'$ is obtained by $\rho$ switching  $\delta_i$ and $\delta_{i+1}$ (which is possible by condition \ref{cond:CR} and deterministic up to iso by condition \ref{cond:CRdet}) then $\delta'_i$ and $\delta'_{i+1}$ are weak sequential independent in $\rho'$.

 Then $\si$ satisfies all the conditions of Definition~\ref{def:si-conditions}.
 \end{proposition}

\begin{proof}
Conditions  \ref{cond:CR} and \ref{cond:CRdet} already hold for weak sequential independence. Condition \ref{cond:si-symm} holds by definition. \textbf{Condition \ref{cond:si-si-iso} should be proved explicitly, I fear...}
\end{proof}

\begin{definition}[Invariance under shift]
\label{def:invariance} 
A binary relation $R$ on consecutive steps of derivations is \emph{invariant under shift equivalence up to $n$} if whenever  $\rho = \rho_0 \shiftdir[\upsilon_1] \rho_1 \ldots \shiftdir[\upsilon_k] \rho'$ with  $\perm = \upsilon_1; \ldots; \upsilon_k$ and $k \leq n$ (thus $\rho \shifteq[\perm] \rho'$ with a proof made of at most  $n$ switches of steps) and $\perm(i+1) = \perm(i) + 1$, then $\delta_i \mathrel{R} \delta_{i+1}$ in $\rho$ if and only if $\delta'_{\perm(i)} \mathrel{R} \delta'_{\perm(i+1)}$ in $\rho'$. Relation $R$  is \emph{invariant under shift equivalence}, tout-court, if it is so up to any $n$.

In words, two consecutive steps are related by $R$ in a derivation iff they are related by $R$ in any shift-equivalent derivation in which they are again consecutive.

%$\delta_i \mathrel{R} \delta_{i+1}$ in $\rho$, $\rho \shifteq[\perm] \rho'$, and $\perm(i+1) = \perm(i) + 1$, then $\delta'_{\perm(i)} \mathrel{R} \delta'_{\perm(i+1)}$ in $\rho'$.

\end{definition}

\todo[inline]{The following is Lemma~\ref{le:no-useless-shifts} with a bound $n$ on the length of the proof that two derivations are shift equivalent. Also, it uses as assumption the fact that \si is invariant w.r.t.~shift up to $n-1$: the fact that $n-1$ is sufficient is exploited later in the proof of the main result. I think that it works, because the assumption is used only at the end, when the two reverse transpositions were already stripped, thus the length is $n-2$.}
\begin{lemma}[no need of useless shifts]
  \label{le:no-useless-shifts-bounded}
  Given an integer $n$, let us assume that $\si$ is invariant under shift up to $n-1$.  Let $\rho$, $\rho'$ be direct derivations such that  $\rho = \rho_0 \shiftdir[\upsilon_1] \rho_1 \ldots \shiftdir[\upsilon_k] \rho'$ with $k \leq n$, and 
  let $\sigma = \upsilon_1; \ldots; \upsilon_k$.
   Then
  $\rho = \rho_0 \shiftdir[\upsilon'_1] \rho_1 \ldots \shiftdir[\upsilon'_h ]
  \rho'$ where $\sigma = \upsilon'_1; \ldots; \upsilon'_h$, $h \leq k$ and $\upsilon'_j \in \inv{\upsilon'_{j,h}}$ for $j \in \{1, \ldots,h\}$. 
  %\todo{I understand that $\upsilon'_{j,h} = \upsilon'_j; \ldots; \upsilon'_h$}
\end{lemma}

\begin{proof}
  Let 
  $\rho = \rho_0 \shiftdir[\upsilon_1] \rho_1 \ldots \shiftdir[\upsilon_k ]
  \rho'$.
  %
  We show that if some $\upsilon_j$ does not act on an inversion, i.e.,
  $\upsilon_j \not\in \inv{\upsilon_{j,k}}$ then we can shorten the
  proof of $\rho \shifteq[\sigma] \rho'$ of two steps. This 
  allow us to conclude that we can reduce the original proof to one which is not longer and only acting
  on inversions, as desired.

  The key observation is that, if
  $\upsilon_j \not\in \inv{\sigma_{j,k}}$ then, intuitively, the ``same
  exchange'' must be performed in the reverse direction by some
  $\upsilon_s$, with $s>j$.  Formally, if $\upsilon_j = \transp{i}$ and
  $(i,i+1)\not\in\inv{\upsilon_{i,k}}$ then there exists $s>j$ such that
  $\upsilon_{j,s-1}(i) = \upsilon_{j,s-1}(i+1)+1$ and
  $\upsilon_s = \transp{\upsilon_{j,s-1}(i+1)}$. Among all the pairs
  $j, s$ of this kind, take one with minimal $s-j$. This means that in
  the middle all transpositions are inversions of $\upsilon_{j+1,s-1}$.

  By Lemma~\ref{le:perm}, $\upsilon_j$ and $\upsilon_s$ can be
  stripped, i.e., $\upsilon_{j,s} = \upsilon_{j+1,s-1}$. Moreover,
  again by Lemma~\ref{le:perm},
  $\upsilon_{i+1}, \ldots, \upsilon_{s-1}$ perform exchanges that were
  already done in the original proof, and thus, since independence is
  invariant under shift equivalence up to $n-1$, they are allowed. 
  Moreover, since by condition \ref{cond:CRdet} of Definition~\ref{def:si-conditions} the result of switching is uniquely
  determined (up to iso), the resulting derivation is the same as
  before, hence we can strip $\upsilon_{j}$ and $\upsilon_{s}$,
  and conclude.
\end{proof}

\todo[inline]{The next one is Lemma~\ref{le:seq-ind-glob2}, again bounded w.r.t.~ the length of the proof of shift equivalence. \textbf{Note:} It is not used in the following.}

Then we can prove that whenever $\rho \shifteq[\perm] \rho'$, each
decomposition of $\perm$ in terms of inversions leads to a valid
valid proof of the shift equivalence  of $\rho$ and $\rho$.

\begin{lemma}[independence is global/2]
  \label{le:seq-ind-glob2-new}
 Given an integer $n$, let us assume that $\si$ is invariant under shift up to $n$.  Let $\rho$, $\rho'$ be direct derivations such that $\rho = \rho_0 \shiftdir[\upsilon_1] \rho_1 \ldots
  \shiftdir[\upsilon_k] \rho'$, with $0 < k \leq n$,  
   $\sigma = \upsilon_1; \ldots; \upsilon_k$  and  $\upsilon_j \in \inv{\upsilon_{j,k}}$ for $j \in \{1, \ldots,k\}$. 
  Let
  $\perm = \transp{i}; \perm'$ where $(i,i+1) \in \inv{\perm}$. Then
  there exists $\rho''$ such that
  $\rho \shifteq[\transp{i}] \rho'' \shifteq[\perm']
  \rho'$.
\end{lemma}

\begin{proof}
%  Let
%  $\rho = \rho_0 \shiftdir[\upsilon_1] \rho_1 \ldots
%  \shiftdir[\upsilon_n] \rho'$ where $\sigma = \upsilon_1; \ldots; \upsilon_n$ and $\upsilon_j \in \inv{\upsilon_{j,n}}$ for $j \in \{1, \ldots,n\}$, be the decomposition
%  provided by Lemma~\ref{le:no-useless-shifts}.
 % 
  We know that steps corresponding to $i$ and $i+1$ are exchanged at
  some point, namely there is $j$ such that
  $\upsilon_j =
  \transp{\upsilon_{1,j-1}(i)}[\upsilon_{1,j-1}(i+1)]$. Since $\si$ is invariant up to $n$, the steps $i$ and $i+1$ are independent
  already at the beginning and thus we can exchange them first.
  More precisely, 
  \begin{center}
    $\rho \shiftdir_{\transp{i}} \rho'' \shiftdir_{\upsilon_{1,j-1}}
    \rho_j \shiftdir_{\upsilon_{j+1,k}} \rho'$.
  \end{center}
  This can be proved by induction on $j$.
\end{proof}

The following main lemma needs a proof.

\begin{lemma}[Invariance of $\si$ under immediate shift]
\label{le:invariance-immediate}
\begin{enumerate}
%\item Let $\rho = \delta_1 ; \delta_2 ; \delta_3 $ with $\delta_2 \si \delta_3$, suppose that $\rho \shiftdir[(2,3)] \rho' = \delta_1 ; \delta'_3 ; \delta'_2$,  that $\delta_1 \si \delta'_3$ in $\rho'$, and that $\rho' \shiftdir[(1,2)] \rho'' =  \delta''_3 ; \delta'_1 ; \delta'_2$. 
%Then $\delta_1 \si \delta_2$ in $\rho$ if and only if  $\delta'_1 \si \delta'_2$ in $\rho''$.

\item Let $\rho = \delta_1 ; \delta_2 ; \delta_3 $ be a derivation such that $\rho \shiftdir[(2,3)] \rho'  \shiftdir[(1,2)] \rho''$,  with $\delta_2 \si \delta_3$ in $\rho$, $\rho' = \delta_1 ; \delta'_3 ; \delta'_2$,   $\delta_1 \si \delta'_3$ in $\rho'$, and that $\rho'' =  \delta''_3 ; \delta'_1 ; \delta'_2$. 
Then $\delta_1 \si \delta_2$ in $\rho$ if and only if  $\delta'_1 \si \delta'_2$ in $\rho''$.

The situation can be represented as in the following diagram, where each transposition is performed, downwards, only if the two starting steps are sequential independent. The proposition states that the steps \cycl{1} are sequential independent if and only if so are the steps \cycl{2}. For brevity, we abbreviate this statement to $SI(\cycl{1}) \iff SI(\cycl{2})$.

%\center{\xymatrix@C=2pc@R=1pc{
% \rho  \ar@{~>}[d]^{sh}_{(2,3)}  \ar@{}[r]|{=} & \rho_1 \ar@{-}[d] \ar@{}[r]|{;} & \delta_1 \ar@{-}[d]  \ar@{}[r]|{;} & \delta_2 \ar@{-}[dr]  \ar@{}[r]|{;} & \delta_3 \ar@{-}[dl]  \ar@{}[r]|{;} &\rho_2 \ar@{-}[d] \\
%\rho'  \ar@{~>}[d]^{sh}_{(1,2)}  \ar@{}[r]|{=} &  \rho_1 \ar@{-}[d]  \ar@{}[r]|{;} & \delta_1 \ar@{-}[dr]  \ar@{}[r]|{;} & \delta'_3 \ar@{-}[dl]  \ar@{}[r]|{;} & \delta'_2 \ar@{-}[d]  \ar@{}[r]|{;} &\rho_2 \ar@{-}[d] \\
%\rho''  \ar@{}[r]|{=} &   \rho_1 \ar@{}[r]|{;} & \delta''_3 \ar@{}[r]|{;} & \delta'_1 \ar@{}[r]|{;} & \delta'_2 \ar@{}[r]|{;} &\rho_2\\
%}}

$$\xymatrix@C=2pc@R=1pc{
 \rho  \ar@{~>}[d]^{sh}_{(2,3)}  \ar@{}[r]|{=} &  \delta_1 \ar@{-}[d]  \ar@{}[r]|{;}  \ar@{-}@/^3ex/[r]|{\cycl{1}} & \delta_2 \ar@{-}[dr]  \ar@{}[r]|{;} & \delta_3 \ar@{-}[dl] \\
\rho'  \ar@{~>}[d]^{sh}_{(1,2)}  \ar@{}[r]|{=} &   \delta_1 \ar@{-}[dr]  \ar@{}[r]|{;} & \delta'_3 \ar@{-}[dl]  \ar@{}[r]|{;} & \delta'_2 \ar@{-}[d]\\
\rho''  \ar@{}[r]|{=} &    \delta''_3 \ar@{}[r]|{;} & \delta'_1 \ar@{-}@/_3ex/[r]|{\cycl{2}} \ar@{}[r]|{;} & \delta'_2  \\
}$$

\item  Also a symmetric statement holds, for which we show only the graphical representation. Again, we show that $SI(\cycl{1}) \iff SI(\cycl{2})$.
% the steps \cycl{1} are sequential independent if and only if so are the steps \cycl{2}.

$$\xymatrix@C=2pc@R=1pc{
 \rho  \ar@{~>}[d]^{sh}_{(1,2)}  \ar@{}[r]|{=} &
   \delta_1 \ar@{-}[dr]  \ar@{}[r]|{;}  &
    \delta_2 \ar@{-}[dl]  \ar@{}[r]|{;} \ar@{-}@/^3ex/[r]|{\cycl{1}} &
    \delta_3 \ar@{-}[d] \\
\rho'  \ar@{~>}[d]^{sh}_{(2,3)}  \ar@{}[r]|{=} &
   \delta'_2 \ar@{-}[d]  \ar@{}[r]|{;} &
    \delta'_1 \ar@{-}[dr]  \ar@{}[r]|{;} &
     \delta_3 \ar@{-}[dl]\\
\rho''  \ar@{}[r]|{=} & 
   \delta'_2 \ar@{}[r]|{;} \ar@{-}@/_3ex/[r]|{\cycl{2}} &
    \delta'_3  \ar@{}[r]|{;} &
     \delta''_1  \\
}$$

%
%Symmetrically, let $\rho = \rho_1 ; \delta_1 ; \delta_2 ; \delta_3 ; \rho_2$ with $\delta_1 \si \delta_2$ in $\rho$, suppose that $\rho \shiftdir[(1,2)] \rho' = \rho_1 ; \delta'_2 ; \delta'_1 ; \delta_3 ;\rho_2$,  that $\delta'_1 \si \delta_3$ in $\rho'$, and that $\rho' \shiftdir[(2,3)] \rho'' = \rho_1 ; \delta'_2 ; \delta'_3 ; \delta''_1 ;\rho_2$. 
%Then $\delta_2 \si \delta_3$ in $\rho$ if and only if  $\delta'_2 \si \delta'_3$ in $\rho''$. 
\end{enumerate}
%\begin{enumerate}
%\item Let $\rho = \rho_1 ; \delta_1 ; \delta_2 ; \delta_3 ; \rho_2$ with $\delta_2 \si \delta_3$ in $\rho$, suppose that $\rho \shiftdir[(2,3)] \rho' = \rho_1 ; \delta_1 ; \delta'_3 ; \delta'_2 ;\rho_2$,  that $\delta_1 \si \delta'_3$ in $\rho'$, and that $\rho' \shiftdir[(1,2)] \rho'' = \rho_1 ; \delta''_3 ; \delta'_1 ; \delta'_2 ;\rho_2$. 
%Then $\delta_1 \si \delta_2$ in $\rho$ if and only if  $\delta'_1 \si \delta'_2$ in $\rho''$. 
%
%The following A graphical representation The situation can 
%\item  Symmetrically, let $\rho = \rho_1 ; \delta_1 ; \delta_2 ; \delta_3 ; \rho_2$ with $\delta_1 \si \delta_2$ in $\rho$, suppose that $\rho \shiftdir[(1,2)] \rho' = \rho_1 ; \delta'_2 ; \delta'_1 ; \delta_3 ;\rho_2$,  that $\delta'_1 \si \delta_3$ in $\rho'$, and that $\rho' \shiftdir[(2,3)] \rho'' = \rho_1 ; \delta'_2 ; \delta'_3 ; \delta''_1 ;\rho_2$. 
%Then $\delta_2 \si \delta_3$ in $\rho$ if and only if  $\delta'_2 \si \delta'_3$ in $\rho''$. 
%\end{enumerate}
\end{lemma}
\begin{proof}
\textbf{To be done...}

\end{proof}

We are now able to state and prove the main result.

\begin{theorem}[Invariance of $\si$]
Relation $\si$ is invariant under shift equivalence, as for Definition~\ref{def:invariance}.
\end{theorem}

\begin{proof}
Let $\rho \shifteq[\perm] \rho'$ be two shift equivalent derivations and let $1 \leq i <  |\rho|$ such that $\perm(i+1) = \perm(i) + 1$. We have to show that $\delta_i \si \delta_{i+1}$ in $\rho$  iff  $\delta_{\perm(i)} \si \delta_{\perm(i+1)}$ in $\rho'$.

Since $\rho \shifteq[\perm] \rho'$, we have $\rho = \rho_0 \shiftdir[\upsilon_1] \rho_1 \ldots \shiftdir[\upsilon_n] \rho'$, with $\sigma = \upsilon_1; \ldots; \upsilon_n$. We proceed by induction on $n$. If $n = 0$, then $\perm = id$ and the statement trivially holds. Now we assume that the statement holds for each $k \leq n$, and  we show that it holds for $n+1$. Let  $\sigma'$ be defined as $\sigma' =  \upsilon_1; \ldots; \upsilon_n$, thus  $\sigma = \sigma' ; \upsilon_{n+1}$.  We proceed by case analysis on the last adjacent transposition $\upsilon_{n+1} = (z, z+1)$.

\begin{description} 
\item [{[$z+1 < \sigma'(i)$ or $\sigma'(i+1) < z$]}] In this case the last transposition does not affect the steps $\delta^n_{\sigma'(i)}$ and $\delta^n_{\sigma'(i+1)}$  of $\rho_n$, as depicted here:

$$\xymatrix@C=2pc@R=2pc{
 \rho  \ar@{~>}[d]^{sh}^(.9){*} \ar@{}[r]|{=} &
   \cdots   \ar@{}[r]|{;}  &
    \delta_i \ar@{--}@/_2ex/[d]  \ar@{}[r]|{;} \ar@{-}@/^3ex/[r]|{\cycl{1}} &
    \delta_{i+1} \ar@{--}@/^1ex/[d]  \ar@{}[r]|{;}  & \cdots\\
\rho_n  \ar@{~>}[d]^{sh}_{\upsilon_{n+1}}  \ar@{}[r]|{=} &
   \cdots   \ar@{}[r]|{;}  &
    \delta^n_{\sigma'(i)} \ar@{-}[d]  \ar@{}[r]|{;}  \ar@{-}@/_3ex/[r]|{\cycl{3}}  &
   \delta^n_{\sigma'(i+1)} \ar@{-}[d]  \ar@{}[r]|{;} &
    \cdots\\
\rho'  \ar@{}[r]|{=} & 
    \cdots   \ar@{}[r]|{;}  &
     \delta'_{\sigma(i)} \ar@{}[r]|{;} \ar@{-}@/_3ex/[r]|{\cycl{2}} &
    \delta'_{\sigma(i+1)}  \ar@{}[r]|{;} &
     \cdots  \\
}$$

\todo[inline]{But is it obvious that if two steps keep their position the remain shift-equivalent? Maybe this is another condition...}

Therefore clearly $SI(\cycl{2}) \iff SI(\cycl{3})$. Furthermore $SI(\cycl{1}) \iff SI(\cycl{3})$  by induction hypothesis, because $\rho  \stackrel{sh\ *}{\rightsquigarrow} \rho_n$  has length $n$, and thus we conclude $SI(\cycl{1}) \iff SI(\cycl{2})$.

\item  [{[$z = \sigma'(i+1)$ and $\sigma'(i) = z+1$]}]  In this case the last transposition switches the steps $\delta^n_{\sigma'(i+1)}$ and $\delta^n_{\sigma'(i)}$  of $\rho_n$, as depicted here:

$$\xymatrix@C=2pc@R=2pc{
 \rho  \ar@{~>}[d]^{sh}^(.9){*} \ar@{}[r]|{=} &
   \cdots   \ar@{}[r]|{;}  &
    \delta_i \ar@{--}[dr]  \ar@{}[r]|{;} \ar@{-}@/^3ex/[r]|{\cycl{1}} &
    \delta_{i+1} \ar@{--}[dl]  \ar@{}[r]|{;}  & \cdots\\
\rho_n  \ar@{~>}[d]^{sh}_{\upsilon_{n+1}}  \ar@{}[r]|{=} &
   \cdots   \ar@{}[r]|{;}  &
    \delta^n_{\sigma'(i+1)} \ar@{-}[dr]  \ar@{}[r]|{;}  \ar@{-}@/^2ex/[r]|{\cycl{3}}  &
   \delta^n_{\sigma'(i)} \ar@{-}[dl]  \ar@{}[r]|{;} &
    \cdots\\
\rho'  \ar@{}[r]|{=} & 
    \cdots   \ar@{}[r]|{;}  &
     \delta'_{\sigma(i)} \ar@{}[r]|{;} \ar@{-}@/_3ex/[r]|{\cycl{2}} &
    \delta'_{\sigma(i+1)}  \ar@{}[r]|{;} &
     \cdots  \\
}$$

Clearly we have $SI(\cycl{3})$ because the two steps are switched, and thus also $SI(\cycl{2})$ by  conditions  \ref{cond:si-symm} and \ref{cond:si-si-iso} of Definition~\ref{def:si-conditions}. Furthermore, as evident from the drawing the steps $\delta_i$ and $\delta_{i+1}$ of $\rho$ must have been switched by a transposition $\upsilon_j = (v, v+1)$ with $1 \leq j \leq n$.  Therefore we have that steps $\delta^{j-1}_v$ and $\delta^{j-1}_{v+1}$  are sequential independent in $\rho^{j-1}$, and by induction hypothesis  (because $j \leq n$) we infer $SI(\cycl{1})$, allowing us to conclude that $SI(\cycl{1}) \iff SI(\cycl{2})$.

\item  [{[$z+1 = \sigma'(i+1)$, thus $z = \sigma(i+1)$]}]

% and $z =\sigma'(v) for $v \gt i+1
In this case the transposition $\upsilon_{n+1}$ affects one of the steps \cycl{2} but not both, as summarized in the drawing.
$$\xymatrix@C=2pc@R=2pc{
 \rho  \ar@{~>}[d]^{sh}^(.9){*} \ar@{}[r]|{=} &
   \cdots   \ar@{}[r]|{;}  &
    \delta_i \ar@{--}@/_2ex/[d]  \ar@{}[r]|{;} \ar@{-}@/^3ex/[r]|{\cycl{1}} &
    \delta_{i+1} \ar@{--}@/^1ex/[dr]  \ar@{}[r]|{;}  & \cdots\\
\rho_n  \ar@{~>}[d]^{sh}_{\upsilon_{n+1}}  \ar@{}[r]|{=} &
   \cdots   \ar@{}[r]|{;}  &
    \delta^n_{\sigma'(i)} \ar@{-}[d]  \ar@{}[r]|{;}   &
   \delta^n_z \ar@{-}[dr]  \ar@{}[r]|{;} &
   \delta^n_{\sigma'(i+1)} \ar@{-}[dl]  \ar@{}[r]|{;} &
    \cdots\\
\rho'  \ar@{}[r]|{=} & 
    \cdots   \ar@{}[r]|{;}  &
     \delta'_{\sigma(i)} \ar@{}[r]|{;} \ar@{-}@/_3ex/[r]|{\cycl{2}} &
    \delta'_{\sigma(i+1)}  \ar@{}[r]|{;} &
    \delta'_{z+1}  \ar@{}[r]|{;} &
     \cdots  \\
}$$

There are two subcases, depending on whether step $\delta_z^n$ originates from a step in $\rho$ after $\delta_{i+1}$ or before $\delta_i$. In the first case we have:  
$$\xymatrix@C=2pc@R=2pc{
 \rho  \ar@{~>}[d]^{sh}^(.9){*} \ar@{}[r]|{=} &
   \cdots   \ar@{}[r]|{;}  &
    \delta_i \ar@{--}@/_2ex/[d]  \ar@{}[r]|{;} \ar@{-}@/^3ex/[r]|{\cycl{1}} &
   \delta_{i+1}  \ar@{--}@/^1ex/[dr]  \ar@{}[r]|{\cdots}  &
    \delta_{{\sigma'}^{-1}(z)} \ar@{--}@/_1ex/[dl]  \ar@{}[r]|{;}  & \cdots\\
\rho_n  \ar@{~>}[d]^{sh}_{\upsilon_{n+1}}  \ar@{}[r]|{=} &
   \cdots   \ar@{}[r]|{;}  &
    \delta^n_{\sigma'(i)} \ar@{-}[d]  \ar@{}[r]|{;}   &
   \delta^n_z \ar@{-}[dr]  \ar@{}[r]|{;} &
   \delta^n_{\sigma'(i+1)} \ar@{-}[dl]  \ar@{}[r]|{;} &
    \cdots\\
\rho'  \ar@{}[r]|{=} & 
    \cdots   \ar@{}[r]|{;}  &
     \delta'_{\sigma(i)} \ar@{}[r]|{;} \ar@{-}@/_3ex/[r]|{\cycl{2}} &
    \delta'_{\sigma(i+1)}  \ar@{}[r]|{;} &
    \delta'_{z+1}  \ar@{}[r]|{;} &
     \cdots  \\
}$$

As evident form the drawing, in this case there are certainly some useless transpositions. Thus by Lemma~\ref{le:no-useless-shifts-bounded} (used on proofs of length at most $n+1$, with the induction hypothesis up to length $n$) there exists an equivalent proof of $\rho \shifteq[\perm] \rho'$  of length at most $n-1$, thus we can conclude by inductive hypothesis.

The second case is depicted as follows:

$$\xymatrix@C=2pc@R=2pc{
 \rho  \ar@{~>}[d]^{sh}^(.9){*} \ar@{}[r]|{=} &
   \delta_{{\sigma'}^{-1}}(z)  \ar@{--}@/^1ex/[drr]   \ar@{}[r]|{\cdots}  &
    \delta_i \ar@{--}@/_2ex/[d]  \ar@{}[r]|{;} \ar@{-}@/^3ex/[r]|{\cycl{1}} &
    \delta_{i+1} \ar@{--}@/^1ex/[dr]  \ar@{}[r]|{;}  & \cdots\\
\rho_n  \ar@{~>}[d]^{sh}_{\upsilon_{n+1}}  \ar@{}[r]|{=} &
   \cdots   \ar@{}[r]|{;}  &
    \delta^n_{\sigma'(i)} \ar@{-}[d]  \ar@{}[r]|{;}   &
   \delta^n_z \ar@{-}[dr]  \ar@{}[r]|{;} &
   \delta^n_{\sigma'(i+1)} \ar@{-}[dl]  \ar@{}[r]|{;} &
    \cdots\\
\rho'  \ar@{}[r]|{=} & 
    \cdots   \ar@{}[r]|{;}  &
     \delta'_{\sigma(i)} \ar@{}[r]|{;} \ar@{-}@/_3ex/[r]|{\cycl{2}} &
    \delta'_{\sigma(i+1)}  \ar@{}[r]|{;} &
    \delta'_{z+1}  \ar@{}[r]|{;} &
     \cdots  \\
}$$

We see that two steps from which $\delta^n_{\sigma'(i)}$ and $\delta^n_z$ originate were switched by a transposition $\upsilon_j$ with $1 \leq j \leq n$ therefore by induction hypothesis we infer that  $\delta^n_{\sigma'(i)}$ and $\delta^n_z$ are sequentially independent. Thus we can modify the proof as follows, switching such steps twice:

$$\xymatrix@C=2pc@R=2pc{
 \rho  \ar@{~>}[d]^{sh}^(.9){*} \ar@{}[r]|{=} &
   \delta_{{\sigma'}^{-1}}(z)  \ar@{--}@/^1ex/[drr]   \ar@{}[r]|{\cdots}  &
    \delta_i \ar@{--}@/_2ex/[d]  \ar@{}[r]|{;} \ar@{-}@/^3ex/[r]|{\cycl{1}} &
    \delta_{i+1} \ar@{--}@/^1ex/[dr]  \ar@{}[r]|{;}  & \cdots\\
\rho_n  \ar@{~>}[d]^{sh}_{\upsilon_{n+1}}  \ar@{}[r]|{=} &
   \cdots   \ar@{}[r]|{;}  &
    \delta^n_{\sigma'(i)} \ar@{-}[dr]  \ar@{}[r]|{;}   &
   \delta^n_z \ar@{-}[dl]  \ar@{}[r]|{;} &
   \delta^n_{\sigma'(i+1)} \ar@{-}[d]  \ar@{}[r]|{;} &
    \cdots\\
  \rho_{n+1}  \ar@{~>}[d]^{sh}_{\upsilon_{n+2}}  \ar@{}[r]|{=} &
   \cdots   \ar@{}[r]|{;}  &
    \delta^{n+1}_{z-1} \ar@{-}[dr]  \ar@{}[r]|{;}   &
   \delta^{n+1}_{\sigma'(i)+1}  \ar@{-}[dl]  \ar@{}[r]|{;}   \ar@{-}@/^3ex/[r]|{\cycl{3}} &
   \delta^{n+1}_{\sigma'(i+1)} \ar@{-}[d]  \ar@{}[r]|{;} &
    \cdots\\
    \rho_{n+2}  \ar@{~>}[d]^{sh}_{\upsilon_{n+3}}  \ar@{}[r]|{=} &
   \cdots   \ar@{}[r]|{;}  &
    \delta^{n+2}_{\sigma'(i)} \ar@{-}[d]  \ar@{}[r]|{;}   &
   \delta^{n+2}_z \ar@{-}[dr]  \ar@{}[r]|{;} &
   \delta^{n+2}_{\sigma'(i+1)} \ar@{-}[dl]  \ar@{}[r]|{;} &
    \cdots\\  
\rho'  \ar@{}[r]|{=} & 
    \cdots   \ar@{}[r]|{;}  &
     \delta'_{\sigma(i)} \ar@{}[r]|{;} \ar@{-}@/_3ex/[r]|{\cycl{2}} &
    \delta'_{\sigma(i+1)}  \ar@{}[r]|{;} &
    \delta'_{z+1}  \ar@{}[r]|{;} &
     \cdots  \\
}$$

Now we can conclude observing that $SI(\cycl{2}) \iff SI(\cycl{3})$ by point 2 of Lemma~\ref{le:invariance-immediate}, and that $SI(\cycl{1}) \iff SI(\cycl{3})$ can be proved by applying the first subcase of this step of the proof.

\todo[inline]{Lemma~\ref{le:invariance-immediate} considera solo derivazioni di lunghezza 3. Anche qui serve un lemma che garantisce che vale anche in un contesto piu' ampio.}

\item   [{[$z = \sigma'(i)$, thus $z+1 = \sigma(i)$]}]  
This case is completely symmetrical to the previous one, exploiting in the last subcase point 1 of Lemma~\ref{le:invariance-immediate}.

\end{description}

\end{proof}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: