### 2024-02-03 icalp/fusione-pratendsec3.tex

\makeatletter\Hy@SaveLastskip\label{proofsection:prAtEndxi}\ifdefined\pratend@current@sectionlike@label\immediate\write\@auxout{\string\gdef\string\pratend@section@for@proofxi{\pratend@current@sectionlike@label}}\fi\Hy@RestoreLastskip\makeatother\begin{proof}[Proof of \pratendRef{thm:prAtEndxi}]\phantomsection\label{proof:prAtEndxi}The first point of \Cref {def:switc} is obviously satisfied. The second one follows at once since \[m_{\der {E},0}= f_0\circ i_1 \qquad h_{\der {E},1}= g_{1}\circ i_0\] Finally, the third point holds because we already know that \[m_{\der {E},0}= f_0\circ i_1 \qquad h_{\der {E},1}= g_{1}\circ i_0\]\end{proof}

\makeatletter\Hy@SaveLastskip\label{proofsection:prAtEndxii}\ifdefined\pratend@current@sectionlike@label\immediate\write\@auxout{\string\gdef\string\pratend@section@for@proofxii{\pratend@current@sectionlike@label}}\fi\Hy@RestoreLastskip\makeatother\begin{proof}[Proof of \pratendRef{thm:prAtEndxii}]\phantomsection\label{proof:prAtEndxii}aaa\end{proof}