### 2024-02-06 icalp/fusione-pratendsec3.tex

\makeatletter\Hy@SaveLastskip\label{proofsection:prAtEndiii}\ifdefined\pratend@current@sectionlike@label\immediate\write\@auxout{\string\gdef\string\pratend@section@for@proofiii{\pratend@current@sectionlike@label}}\fi\Hy@RestoreLastskip\makeatother\begin{proof}[Proof of \pratendRef{thm:prAtEndiii}]\phantomsection\label{proof:prAtEndiii}Let us begin considering the diagram built in \Cref {rem:deco}: \[\xymatrix {&&R_0 \ar @/_1cm/[ddrr]_(.35){i_0}|(.7)\hole \ar [dr]^{h_0}&& L_1\ar @/^1cm/[ddll]^(.35){i_1} \ar [dl]_{m_1}\\&K_0\ar [dr]^{k_0}\ar @{>->}[dl]_{l_0} \ar @/_.8cm/[ddrr]^(.65){u_0}\ar [ur]^{r_0}&& G_1 && K_1 \ar @/^.8cm/[ddll]_(.65){u_1}\ar [dl]_{k_1}\ar @{>->}[lu]_{l_1} \ar [dr]^{r_1}\\L_0 \ar [dr]^{m_0} \ar @{>}@/_.8cm/[ddrr]_{j_1}&& D_0 \ar @{>->}[dl]|(.42)\hole _(.64){f_0}\ar [ur]|(.48)\hole ^(.7){g_1}&&D_1 \ar [dr]|(.42)\hole ^(.65){g_1} \ar @{>->}[ul]|(.48)\hole _(.7){f_1}&&R_1\ar @/^.8cm/[ddll]^{j_0}\ar [dl]^{h_1}\\&G_0 &&P\ar [dr]^{q_1} \ar @{>->}[dl]_{q_0}\ar [ur]^(.4){p_1}\ar @{>->}[ul]_(.4){p_0}&&G_2\\&&Q_0 \ar @{>->}@{>->}[ul]_{a_0} \ar @{.>}[dr]_{b_0} &&Q_1\ar [ur]^{b_1} \ar @{>.>}[dl]^{a_1} \\ &&&H_1 }\] Since $q_0\colon P\to Q_0$ is in $\mathcal {M}$, we can take its pushout along $q_1$ to get the dotted arrows $a_1\colon Q_1\to H_1$ and $q_0\colon Q_0\to H_1$. But then, since the composite of two pushout squares is a pushout, the diagram below is a derivation. \[\xymatrix @C=15pt{L_1 \ar [d]_{f_0 \circ i_1}&& K_1 \ar [d]_{q_0\circ u_1}\ar @{>->}[ll]_{l_1} \ar [r]^{r_1} & R_1 \ar @/^.35cm/[drrr]|(.3)\hole _(.4){j_0} \ar [dr]|(.3)\hole _{a_1\circ j_0} && L_0 \ar @/_.35cm/[dlll]^(.4){j_1} \ar [dl]|(.3)\hole ^{b_0\circ j_1}& K_1 \ar [d]^{q_1\circ u_0}\ar @{>->}[l]_{l_0} \ar [rr]^{r_0} && R_0 \ar [d]^{g_1\circ i_0} \\G_0 && \ar @{>->}[ll]^{a_0} D_0 \ar [rr]_{b_0}&& G_1 && \ar @{>->}[ll]^{a_1} D_1 \ar [rr]_{b_1}&& G_2}\] It is immediate now to see that it is a switch of $\dder {D}_0$ and $\dder {D}_1$ along $(i_0, i_1)$.\end{proof}