### 2024-07-02 concur/submitted/Figure/coffee-rewriting.tex

\begin{center}

  %%% REWRITING 
  % 
  %
  % 
  \begin{tikzpicture}[node distance=4mm, font=\small]
    % graph G0
    \node (G0) {
      \begin{tikzpicture}[node distance=2mm, font=\scriptsize]
        % 
        \node at (0,0.5) {};
        \node at (0,1.5) [node, label=left:$1$] (1) {} ;
        \node at (0,1.0) [node, label=left:$2$] (2) {} ;
        \node at (0,0.5) [node, label=left:$3$] (3) {} ;

        \node at (0.6,1.0) [node, label=right:$4$] (4) {}
        edge [in=45, out=75, loop] node[above] {$\mathit{cf}$} ();

        \draw[->] (1) to[out=350, in=110] node[above]{$m$} (4);
        \draw[->] (2) to node[above, pos=0.4]{$s$} (4);
        \draw[->] (3) to[out=10, in=240] node[below]{$c$} (4);

      %
      \pgfBox
      \end{tikzpicture} 
    };
    \node[below] at (G0.south) {$G_0$};
    % graph G1
    \node [right=of G0] (G1) {
      \begin{tikzpicture}[node distance=2mm, font=\scriptsize]
      % 
        \node at (0,0.5) {};
        \node at (0,0.5) [node, label=left:$3$] (3) {} ;

        \node at (0.6,1.0) [node, label=right:$124$] (124) {}
        edge [in=90, out=130, loop] node[above] {$\mathit{m}$} ()
        edge [in=155, out=185, loop] node[above] {$\mathit{s}$} ()

        edge [in=45, out=75, loop] node[above] {$\mathit{cf}$} ();

        \draw[->] (3) to[out=10, in=240] node[below]{$c$}  (124);

      %
      \pgfBox
      \end{tikzpicture}
    };
    \node[below] at (G1.south) {$G_1$};

    % graph G2
    \node [right=of G1] (G2) {        
      \begin{tikzpicture}[node distance=2mm, font=\scriptsize]
      % 
        \node at (0,0.5) {};

        \node at (0.6,1.0) [node, label=right:$1234$] (1234) {}
        edge [in=210, out=240, loop] node[below] {$\mathit{c}$} ()
        edge [in=90, out=130, loop] node[above] {$\mathit{m}$} ()
        edge [in=155, out=185, loop] node[above] {$\mathit{s}$} ()
        edge [in=45, out=75, loop] node[above] {$\mathit{cf}$} ();

      %
      \pgfBox
    \end{tikzpicture}
    };
    \node[below] at (G2.south) {$G_2$};

    % graph G3
    \node [right=of G2] (G3) {
      \begin{tikzpicture}[node distance=2mm, font=\scriptsize]
      % 
        \node at (0,0.5) {};

        \node at (0.6,1.0) [node, label=right:$1234$] (1234) {}
        edge [in=210, out=240, loop] node[below] {$\mathit{c}$} ()
        edge [in=90, out=130, loop] node[above] {$\mathit{m}$} ()
        edge [in=155, out=185, loop] node[above] {$\mathit{s}$} ()
        % edge [in=2500, out=295, loop] node[above]
        edge [in=280, out=310, loop] node[right] {$\mathit{r}$} ()
        edge [in=45, out=75, loop] node[above] {$\mathit{cf}$} ();

        % 
        \pgfBox
      \end{tikzpicture}
    };
    \node[below] at (G3.south) {$G_3$};

    \path (G0) edge[->] node[trans, above] {$\rho_m$} (G1);
    \path (G1) edge[->] node[trans, above] {$\rho_c$} (G2);
    \path (G2) edge[->] node[trans, above] {$\rho_r$} (G3);    
  \end{tikzpicture}
\end{center}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../fusione"
%%% End: