### 2024-09-05 concur/presentazione/pres_lunga.tex
fatto

\documentclass[usenames,dvipsnames]{beamer}
\usepackage[utf8]{inputenc}
\usepackage{stmaryrd}
\usepackage{verbatim}

\usetheme{Padova}

\usepackage[utf8]{inputenc}
\usepackage{graphics}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{mathpartir}
\usepackage{mathrsfs}
\usepackage[all]{xy}
\usepackage{amsfonts}

\usepackage[mathscr]{eucal}
\usepackage{ragged2e}
\addtobeamertemplate{theorem begin}{}{\justifying}
\addtobeamertemplate{block begin}{}{\justifying}
\addtobeamertemplate{exampleblock begin}{}{\justifying}
\addtobeamertemplate{alertblock begin}{}{\justifying}
%%% Bibliography
\usepackage[style=authoryear,backend=bibtex]{biblatex}
\addbibresource{bibliog.bib}

\newcommand{\terms}[1]{\mathcal{{#1}}\text{-}\mathsf{Terms}}
\makeatletter
\renewcommand{\itemize}[1][]{%
	\beamer@ifempty{#1}{}{\def\beamer@defaultospec{#1}}%
	\ifnum \@itemdepth >2\relax\@toodeep\else
	\advance\@itemdepth\@ne
	\beamer@computepref\@itemdepth% sets \beameritemnestingprefix
	\usebeamerfont{itemize/enumerate \beameritemnestingprefix body}%
	\usebeamercolor[fg]{itemize/enumerate \beameritemnestingprefix body}%
	\usebeamertemplate{itemize/enumerate \beameritemnestingprefix body begin}%
	\list
	{\usebeamertemplate{itemize \beameritemnestingprefix item}}
	{\def\makelabel##1{%
			{%
				\hss\llap{{%
						\usebeamerfont*{itemize \beameritemnestingprefix item}%
						\usebeamercolor[fg]{itemize \beameritemnestingprefix item}##1}}%
			}%
		}%
	}
	\fi%
	\beamer@cramped%
	\justifying% NEW
	%\raggedright% ORIGINAL
	\beamer@firstlineitemizeunskip%
}
\makeatother

% Author names in publication list are consistent 
% i.e. name1 surname1, name2 surname2
% See https://tex.stackexchange.com/questions/106914/biblatex-does-not-reverse-the-first-and-last-names-of-the-second-author
%\DeclareNameAlias{author}{first-last}

%%% Suppress biblatex annoying warning
\usepackage{silence}
\WarningFilter{biblatex}{Patching footnotes failed}
\usepackage{physics}
%%% Some useful commands
% pdf-friendly newline in links
\newcommand{\pdfnewline}{\texorpdfstring{\newline}{ }} 
% Fill the vertical space in a slide (to put text at the bottom)

% direct shift
\newcommand{\shift}[1]{\ensuremath{\mathrel{{\leftrightsquigarrow}_{#1}}}}

% shift equivalence
\newcommand{\shifteq}[1][]{\ensuremath{\mathrel{{\equiv}^\mathit{sh}_{#1}}}}

\newcommand{\dder}[1]{\mathscr{#1}}
\usepackage{ragged2e}
\addtobeamertemplate{theorem begin}{}{\justifying}
\addtobeamertemplate{block begin}{}{\justifying}
\addtobeamertemplate{exampleblock begin}{}{\justifying}
\addtobeamertemplate{alertblock begin}{}{\justifying}
\addtobeamertemplate{block alerted begin}{}{\justifying}

\usepackage{etoolbox}
\apptocmd{\thebibliography}{\justifying}{}{}

\usepackage{mathpartir}
\usepackage[usestackEOL]{stackengine}
\usepackage{mathtools,amssymb}
%\newcommand{\ar}[0]{\mathsf{ar}}
\newcommand{\comma}[2]{#1\hspace{1pt} {\downarrow}\hspace{1pt} #2}
\newcommand{\ex}[2]{\mathsf{m}\qty(#1, #2)}
\newcommand{\ext}[2]{\mathsf{m}\qty(#1, #2)}
\newcommand{\cri}[1]{\mathsf{cri}\hspace{-.4pt}\left({#1}\right)}
\newcommand\functorop[1][l]{\csname#1functor\endcsname}
\newcommand\lfunctorop[3]{%
	\setbox0=\hbox{$#2$}%
	\kern\wd0%
	\ensurestackMath{\Centerstack[c]{#1\\ \mathllap{#2\;\,}\mathclap{\DownArrow}\\#3}}%
}
\newcommand\rfunctorop[3]{%
	\setbox0=\hbox{$#2$}%
	\ensurestackMath{\Centerstack[c]{#1\\\mathclap{\UpArrow}\mathrlap{\,\;#2}\\#3}}%
	\kern\wd0%
}
\newcommand\functoropmapsto{\mathrel{\ensurestackMath{\Centerstack[c]{\longmapsto\\ \\\longmapsto}}}}
\setstackgap{L}{1.3\normalbaselineskip}
\newcommand\UpArrow{\rotatebox[origin=c]{90}{$\longrightarrow$\,}}
\newcommand\DownArrow{\rotatebox[origin=c]{-90}{$\longrightarrow$\,}}

\newcommand{\arro}[0]{\mathrel{\rotatebox[origin=c]{270}{$\twoheadrightarrow$}}}

\newcommand{\arr}[2]{{#1}{\arro}{\catname{#2}}}
\newcommand\functor[1][l]{\csname#1functor\endcsname}

\newcommand{\tg}[0]{\catname{TG}_{\Sigma}}
\newcommand{\ari}[0]{\mathsf{ar}}
\newcommand{\lgt}[0]{\mathsf{length}}

\newcommand{\cma}[2]{\mathcal{#1}\hspace{1pt} {\downarrow}\hspace{1pt} \mathcal{#2}}
\newcommand\lfunctor[3]{%
	\setbox0=\hbox{$#2$}%
	\kern\wd0%
	\ensurestackMath{\Centerstack[c]{#1\\ \mathllap{#2\;\,}\mathclap{\DownArrow}\\#3}}%
}
\newcommand\rfunctor[3]{%
	\setbox0=\hbox{$#2$}%
	\ensurestackMath{\Centerstack[c]{#1\\\mathclap{\DownArrow}\mathrlap{\,\;#2}\\#3}}%
	\kern\wd0%
}
\newcommand\functormapsto{\mathrel{\ensurestackMath{\Centerstack[c]{\longmapsto\\ \\\longmapsto}}}}
\setstackgap{L}{1.3\normalbaselineskip}
%\newcommand\DownArrow{\rotatebox[origin=c]{-90}{$\longrightarrow$\,}}

\newcommand{\framefill}{\vskip0pt plus 1filll}
\newcommand{\vin}{\rotatebox[origin=c]{-90}{$\dashv$}}
\newcommand{\catname}[1]{{\normalfont\textbf{#1}}}
\newcommand{\homodd}[0]{\catname{HoMod}}
\newcommand{\jp}[0]{\mathfrak{j}_{\mathcal{M}, \mathcal{N}}}
\newcommand{\slice}[2]{(\catname{#1}\downarrow{#2})}
\newcommand{\supp}[1]{\mathsf{supp}({#1},\mu_{#1})}
%\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
%\newcommand{\abs}[1]{\left\lvert#1\right\rvert}
\newcommand{\nat}{\mathsf{nat}}
\newcommand{\y}{\mathsf{y}}
\newcommand{\im}[1]{\mathrm{Im}(#1)}
\newcommand{\chain}[1]{\catname{Ch}(\catname{#1})}
\newcommand{\homo}[1]{\mathcal{K}(\catname{#1})}
\newcommand{\colim}[0]{\mathrm{colim}}
\newcommand{\rela}[2]{\mathscr{Rel}_\mathcal{#2}(\catname{#1})}
\newcommand{\sub}{\mathsf{Sub}}
\newcommand{\dom}{\mathrm{dom}}
\newcommand{\cod}{\mathrm{cod}}
\newcommand{\co}[1]{\{\abs{#1}\}}
\newcommand{\eq}[3]{[#1 = #2]_{#3}}
\newcommand{\equ}[2]{[#1 \equiv #2]}
\newcommand{\rel}[3]{\frak{R}_{\mathcal{#3}}(#1, #2)}
\newcommand{\sat}[0]{\mathsf{cl}}
\newcommand{\cov}{\vartriangleleft}
\newcommand{\pos}[1]{\mathsf{Pos}_{\mathfrak{#1}}}
\newcommand{\rk}[2]{\mathrm{rk}_{#1}({#2})}
\newcommand{\lind}[1]{\mathcal{L}({#1})}
\newcommand{\class}[1]{\catname{Cl}({#1})}
\newcommand{\hoclass}[1]{\catname{HoCl}({#1})}
\newcommand{\propo}[0]{\mathsf{Prop}}
\newcommand{\term}[2]{\mathbf{Term}({#1}, {#2})}
\newcommand{\formu}[1]{\mathbf{Form}_{#1}}
%\newcommand{\ex}[1]{\mathscr{ex(#1)}}
\newcommand{\inte}[1]{\llbracket{#1}\rrbracket}
\newcommand{\typ}[1]{\mathsf{Type}({#1})}
\newcommand{\Set}{\textbf{\textup{Set}}}
\newcommand{\m}[1]{\textbf{\textup{#1}}}
\newcommand{\sh}{\textbf{\textup{Sh}}}
\NewDocumentCommand{\transp}{m o}{%
	\ensuremath{({#1},%
		\IfNoValueTF{#2}%
		{{#1}+1}%
		{#2}%
		)}
}

\newcommand{\inv}[1]{\mathsf{inv}({#1})}

\newcommand{\modd}[0]{\catname{Mod}}
\newcommand{\des}[1]{Des({#1})}
\newcommand{\prof}[2]{\mathsf{Proof}({#1},{#2})}
\newcommand{\ab}[0]{=-Adj}
\newcommand{\ba}[0]{=-E}
\newcommand{\su}[2]{\mathfrak{s}_{#1}^{#2}}
\newcommand{\ipos}[1]{\catname{Pos}(\catname{#1})}
\newcommand{\Su}[2]{\mathfrak{S}_{#1}^{#2}}
\newcommand{\Pro}[2]{\mathfrak{P}_{#1}^{#2}}
\newcommand{\surr}[0]{\mathsf{surr}}
\newcommand{\propag}[0]{\mathsf{propag}}
\newcommand{\mdl}[1]{\catname{Mod}(#1)}
\newcommand{\alg}[1]{\catname{FAlg}(#1)}
\newcommand{\eim}[1]{\catname{EM}(\m{#1})}
\def\H{\textbf {\textup{H}}}
\def\X{\textbf {\textup{X}}}
\def\Y{\textbf {\textup{Y}}}
\newcommand{\fuz}[1]{\textbf{\textup{Fuz}}({\textbf {\textup{#1}}})}
%\newcommand{\yo}{\mathcal{Y}}
\renewcommand*{\bibfont}{\footnotesize}

\newcommand{\der}[1]{\underline{\dder{#1}}}

\DeclareFontFamily{U}{dmjhira}{}
\DeclareFontShape{U}{dmjhira}{m}{n}{
	<-> s*[0.95] dmjhira
}{}
\DeclareFontSubstitution{U}{dmjhira}{m}{n}

\newcommand{\yo}{%
	\raisebox{0.2\depth}[\fontcharht\font`A][0.1\depth]{%
		\usefont{U}{dmjhira}{m}{n}\symbol{"48}%
	}%
}

\usepackage{tikz-cd}
\usepackage{tikz}
\usetikzlibrary{arrows,shapes,snakes,automata,backgrounds,petri,fit,positioning,calc}
\tikzstyle{node}=[circle, draw=black, minimum size=1mm, inner sep=1.5pt, font=\tiny]

\tikzstyle{trans}=[font=\scriptsize]
\tikzstyle{lab}=[font=\small]
\tikzset{
	coloredge/.style={
		->,
		color=red
		%densely dotted
	}
}

\tikzset{
	colorloop/.style={
		loop,
		color=red
		%densely dotted
	}
}

\tikzcdset{
	perm/.style={
		shorten >=-1mm, shorten <=-1mm,
		-,
		% dotted
	}
}

\newcommand{\pgfBox}{
	\begin{pgfonlayer}{background} 
		\fill[blue!2,thick,draw=black!50,rounded corners,inner sep=3mm] ([xshift=-1.5pt,yshift=-1.5pt]current bounding box.south west) rectangle ([xshift=1.5pt,yshift=1.5pt]current bounding box.north east);
	\end{pgfonlayer}
}

\newcommand{\id}[1]{\mathsf{id}_{#1}}

\title{Left-Linear Rewriting in Adhesive Categories}

\date[May 1977]{Calgary, September 11, 2024}
\begin{document}
	\maketitle

\begin{frame}{Concurrency}\justifying

	Key  idea of concurrency theory:  the behaviour of a computational
	device can be modelled by abstracting its sequences of steps via a suitable equivalence.

	\pause 
	\medskip 
	The equivalence captures when steps are causally unrelated and thus could be
	executed in any order, and possibly in parallel.

	\pause 
	\medskip 
	Defining this equivalence often boils down to having a notion
	of independence between two consecutive steps.

\end{frame}

\begin{frame}{DPO rewriting sytems}\justifying

		We will focus on the \emph{double pushout} (DPO) approach  to  graph rewriting (cfr. \cite{ehrig2006fundamentals}).

	\smallskip 
	\begin{overprint}
		\onslide<1|handout:1>
		A \emph{rewriting rule} is a pair  of arrows $l:K\to L$ and $r:K\to R$ with the same domain. To rewrite an object $G$ according to a rule we proceed in three steps.

		\onslide<2| handout:0>

		A \emph{rewriting rule} is a pair  of arrows $l:K\to L$ and $r:K\to R$ with the same domain. To rewrite an object $G$ according to a rule we proceed in three steps.

		First: find a \emph{match} $m:L\to G$.
		\begin{center}
			\begin{tikzpicture}
				\node(L)at(0,0){$L$};
				\node(K)at(2,0){$K$};
				\node(R)at(4,0){$R$};
				\node(G)at(0,-2){$G$};
				\draw[->](K)--(L)node[pos=0.5, above]{$l$};
				\draw[->](K)--(R)node[pos=0.5, above]{$r$};
				\draw[->, dotted](L)--(G)node[pos=0.5, left]{$m$};
			\end{tikzpicture}
		\end{center}

		\onslide<3| handout:0>

A \emph{rewriting rule} is a pair  of arrows $l:K\to L$ and $r:K\to R$ with the same domain. To rewrite an object $G$ according to a rule we proceed in three steps.

		Second: remove the image of $m$  building a pushout square. \vspace{-0.075cm}
		\begin{center}
			\begin{tikzpicture}
				\node(L)at(0,0){$L$};
				\node(K)at(2,0){$K$};
				\node(R)at(4,0){$R$};
				\node(G)at(0,-2){$G$};
				\node(C)at(2,-2){$C$};
				\draw[->](K)--(L)node[pos=0.5, above]{$l$};
				\draw[->](K)--(R)node[pos=0.5, above]{$r$};
				\draw[->](L)--(G)node[pos=0.5, left]{$m$};
				\draw[->, dotted](K)--(C)node[pos=0.5, right]{$p$};
				\draw[->, dotted](C)--(G)node[pos=0.5, below]{$l'$};
			\end{tikzpicture}
		\end{center}		
		\onslide<4| handout:0>

		A \emph{rewriting rule} in a category $\X$ is a pair  of arrows $l:K\to L$ and $r:K\to R$ with the same domain. To rewrite an object $G$ according to a rule we proceed in three steps.

		Third: fill the hole so obtained, with $R$ taking a pushout.\vspace{-0.075cm}
		\begin{center}\hspace{0.045cm}
			\begin{tikzpicture}
				\node(L)at(0,0){$L$};
				\node(K)at(2,0){$K$};
				\node(R)at(4,0){$R$};
				\node(G)at(0,-2){$G$};
				\node(C)at(2,-2){$C$};
				\node(H)at(4,-2){$H$};
				\draw[->](K)--(L)node[pos=0.5, above]{$l$};
				\draw[->](K)--(R)node[pos=0.5, above]{$r$};
				\draw[->](L)--(G)node[pos=0.5, left]{$m$};
				\draw[->](K)--(C)node[pos=0.5, right]{$p$};
				\draw[->](C)--(G)node[pos=0.5, below]{$l'$};
				\draw[->, dotted ](C)--(H)node[pos=0.5, below]{$r'$};
				\draw[->, dotted ](R)--(H)node[pos=0.5, right]{$q$};
			\end{tikzpicture}
		\end{center}		
	\end{overprint}  	
\end{frame}

\begin{frame}{DPO rewriting system}\justifying
The categorical framework is given by \emph{$\mathcal{M}$-adhesive categories}. \pause 
\begin{definition}[\cite{lack2005adhesive,azzi2019essence,ehrig2014adhesive}]
	Let $\mathcal{M}$ be a class of monos in a category $\X$, containing all isomorphisms, stable under pullbacks and composition. $\X$ is \emph{$\mathcal{M}$-adhesive} if, given a cube below with the hooked arrows in $\mathcal{M}$, the bottom face a pushout and the back faces pullbacks, we have:

	\parbox{5cm}{\centering \begin{tikzpicture}
		\node(C)at(-1,0.4){$C$};
		\node(A)at(1,0.8){$A$};
		\node(B)at(2,0.4){$B$};
		\node(D)at(0,0){$D$};
		\node(A')at(1,2){$A'$};
		\node(B')at(2,1.6){$B'$};
		\node(C')at(-1,1.6){$C'$};
		\node(D')at(0,1.2){$D'$};
		\draw[<-](B')--(A');
		\draw[>->](B')--(B);
		\draw[>->](C')--(C); 
		\draw[>->](D')--(D);
		\draw[->](A')--(C'); 
		\draw[->](B')--(D');
		\draw[->](C')--(D');
		\draw[->](A)--(B);
		\draw[->](C)--(D);
		\draw[->](B)--(D);
		\draw[>-](A')--(1,1.45);
		\draw[->](1,1.35)--(A);
		\draw[>-](A)--(0.05, 0.61);
		\draw[<-](C)--(-0.05,0.59);
\end{tikzpicture}} \quad 	\parbox{4cm}{\centering
	Top face is a pushout \\ $\iff$ \\ Front faces are pullbacks}

\end{definition}
\end{frame}

\begin{frame}{DPO rewriting systems}\justifying
		\begin{definition}
		A \emph{DPO-rewriting system} is a pair $(\X, R)$ where $\X$ is an $\mathcal{M}$-adhesive category and $R$ is a set of rewriting rules.
		\pause

		A system is \emph{linear} if every rule is made by arrows in $\mathcal{M}$, \emph{left linear} if the left hand side of every rule is in $\mathcal{M}$.
	\end{definition}

\pause

In left linear rewriting systems a rule can ``merge'' parts of the state.

\pause 
\begin{alertblock}{Motivation}
	Our long term aim is to provide a concurrent semantics for left linear DPO-rewriting systems (see \cite{baldan2017domains}). 
\end{alertblock}

\end{frame}

\begin{frame}{An example}\justifying

	We want to model a situation in which a coffee comes
	with a bag of sugar, a portion of milk and bag of cocoa.

	\pause 
\begin{figure}
	\input{Figure/coffee-rules2.tex}

\end{figure}
\end{frame}

\begin{frame}{An example}
	If we want a (very sweet) coffee we can add first the milk and sugar, then cocoa. Finally we can stir it.
	\pause 
\begin{figure}
	\input{Figure/coffee-rewriting2.tex}
\end{figure}\pause

$\rho_s$ only needs sugar, so it can be anticipated before $\rho_c$, but in order to do so the presence of $\rho_m$ at the beginning is essential.
\end{frame}

\begin{frame}{Sequential Independence}\justifying

	\begin{definition}  Two consecutive steps are \emph{sequentially independent} it the dotted arrows in the diagram below exist (\emph{independence pair}).
		\begin{center}
			\begin{tikzpicture}
				\node (A)at(0,1){$L_1$};
				\node (B)at(1,1){$K_1$};
				\node (C)at(2,1){$R_1$};
				\node (D)at(4,1){$L_2$};
				\node (E)at(5,1){$K_2$};
				\node(F) at(6,1){$R_2$};
				\node (G)at(0,0){$G_1$};
				\node (H)at(1,0){$H_1$};
				\node (I)at(3,0){$G_2$};
				\node (L)at(5,0){$H_2$};
				\node (M)at(6,0){$G_3$};

				\draw[>->](B)--(A);
				\draw[->](B)--(C);
				\draw[>->](E)--(D);
				\draw[->](E)--(F);
				\draw[>->](H)--(G);
				\draw[->](H)--(I);
				\draw[>->](L)--(I);
				\draw[->](L)--(M);
				\draw[->](A)--(G);
				\draw[->](B)--(H);
				\draw[->](C)--(I);
				\draw[->](D)--(I);
				\draw[->](E)--(L);
				\draw[->](F)--(M);
				\draw[dotted, ->](C) to [out=0, in=150](L);
				\draw[dotted, ->](D)to [out=180, in=30](H);
			\end{tikzpicture}
		\end{center}
		\end{definition}

\pause 
If a system is linear:
\begin{enumerate}
	\item uniqueness of independence pair; \pause 
	\item Local Church-Rosser Theorem: sequentially independent steps can be switched.
\end{enumerate}

\end{frame}

\begin{frame}{Church-Rosser can fail}\justifying

\begin{columns} 
	\begin{column}{.30\textwidth}\centering 	
\xymatrix@C=16pt@R=1pt{
	& 0 \ar@{-}[dd] \ar@{-}[ddddddl] \ar@{-}[ddddddr] &  \\
	&  & \\
	& 1 \ar@{-}[dd] \ar@{-}[ddddl]  \ar@{-}[ddddr]   &  \\
	&  & \\
	& 2 \ar@{.}[d] \ar@{-}[ddl]   \ar@{-}[ddr]     &  \\
	&  \ar@{.}[dl]   \ar@{.}[dr] & \\
	b \ar@{-}[dr] & & c \ar@{-}[dl]\\
	& a } \\
	$\mathcal{M}$ is the class of the identities.	
\end{column}

	\begin{column}{.70\textwidth} \centering
\begin{tikzpicture}
		\node (A)at(0,1){$a$};
		\node (B)at(1,1){$a$};
		\node (C)at(2,1){$0$};
		\node (D)at(4,1){$a$};
		\node (E)at(5,1){$a$};
		\node(F) at(6,1){$b$};
		\node (G)at(0,0){$c$};
		\node (H)at(1,0){$c$};
		\node (I)at(3,0){$0$};
		\node (L)at(5,0){$0$};
		\node (M)at(6,0){$0$};

		\draw[>->](B)--(A);
	\draw[->](B)--(C);
	\draw[>->](E)--(D);
	\draw[->](E)--(F);
	\draw[>->](H)--(G);
	\draw[->](H)--(I);
	\draw[>->](L)--(I);
	\draw[>->](L)--(M);
	\draw[->](A)--(G);
	\draw[->](B)--(H);
	\draw[>->](C)--(I);
	\draw[->](D)--(I);
	\draw[->](E)--(L);
	\draw[->](F)--(M);
	\draw[>->](C) to [out=0, in=150](L);
	\draw[->](D)to [out=180, in=30](H);
	\end{tikzpicture}
\\	
Two rewriting steps.
	\end{column}
\end{columns}

 \centering
 \begin{tikzpicture}
 		\node (A)at(0,1){$a$};
 		\node (B)at(1,1){$a$};
 		\node (C)at(2,1){$b$};

 		\node (G)at(0,0){$c$};
 		\node (H)at(1,0){$c$};

 		\draw[>->](B)--(A);
 		\draw[->](B)--(C);

 		\draw[>->](H)--(G);

 		\draw[->](A)--(G);
 		\draw[->](B)--(H);

 	\end{tikzpicture}
\\
No rewriting step.

\end{frame}

\begin{frame}{Well switching rewriting systems}\justifying

\begin{definition}
A rewriting system $(\X, R)$ is \emph{well switching} of independence pairs are unique and the Local Church-Rosser Theorem holds (sequentially independent derivation can be switched).
\end{definition}
\pause 
\begin{block}{Remark}
	It is possible to identify a larce class of categories on which every left-linear rewriting system enjoys the Local-Church Rosser Theorem holds (\cite{baldan2011adhesivity}).
\end{block}
\pause 
\begin{block}{Example}In the category $\catname{Graph}$, if the rules of a left-linear rewriting system only merge nodes and their left-hand sides contain no isolated nodes, then the system is well-switching.
\end{block}

\end{frame}

\begin{frame}{Results: the Three Steps Lemma}\justifying

\begin{lemma}  Let
	$\dder{D}_0\cdot \dder{D}_1 \cdot \dder{D}_2$ be a three-steps
	derivation in a well-switching rewriting system. 
	\begin{enumerate}\justifying
		\item
		If $\dder{D}_0$ and $\dder{D}_1$ are independent and there is a
		 sequence of switches
		$\dder{D}_0\cdot \dder{D}_1 \cdot \dder{D}_2 \shift{(1,2)}
		\dder{D}_0 \cdot \dder{D}_2' \cdot \dder{D}_1' \shift{(0,1)}
		\dder{D}_2'' \cdot \dder{D}_0' \cdot \dder{D}_1'$ then
		in the last derivation $\dder{D}_0'$ and $\dder{D}_1'$ are independent;

		\item
		\label{lem:indep-global-left:2}
		Suppose that $\dder{D}_1$ and $\dder{D}_2$ are independent, and that, in $\dder{D}_0\cdot \dder{D}_1 \cdot \dder{D}_2 \shift{(1,2)}
		\dder{D}_0 \cdot \dder{D}_2' \cdot \dder{D}_1'$, $\dder{D}_0$ and $\dder{D}_2'$ are independent.  If there
		is a switching sequence
		$\dder{D}_0\cdot \dder{D}_1 \cdot \dder{D}_2 \shift{(0,1)}
		\dder{D}_1' \cdot \dder{D}_0' \cdot \dder{D}_2 \shift{(1,2)}
		\dder{D}_1' \cdot \dder{D}_2' \cdot \dder{D}_0''$ then in the last derivation $\dder{D}_1'$ and $\dder{D}_2'$ are independent.
	\end{enumerate}
\end{lemma}
\end{frame}

\begin{frame}{Results: the Three Steps Lemma}\justifying

	\begin{lemma}  Let
		$\dder{D}_0\cdot \dder{D}_1 \cdot \dder{D}_2$ be a three-steps
		derivation in a well-switching rewriting system. \pause 
		\begin{enumerate}\justifying
			\item
			If $\dder{D}_0$ and $\dder{D}_1$ are independent and there is a
			sequence of switches
			$\dder{D}_0\cdot \dder{D}_1 \cdot \dder{D}_2 \shift{(1,2)}
			\dder{D}_0 \cdot \dder{D}_2' \cdot \dder{D}_1' \shift{(0,1)}
			\dder{D}_2'' \cdot \dder{D}_0' \cdot \dder{D}_1'$ then
			in the last derivation $\dder{D}_0'$ and $\dder{D}_1'$ are independent; \pause

			\item 
			Suppose that $\dder{D}_1$ and $\dder{D}_2$ are independent, and that, in $\dder{D}_0\cdot \dder{D}_1 \cdot \dder{D}_2 \shift{(1,2)}
			\dder{D}_0 \cdot \dder{D}_2' \cdot \dder{D}_1'$, $\dder{D}_0$ and $\dder{D}_2'$ are independent.  If there
			is a switching sequence
			$\dder{D}_0\cdot \dder{D}_1 \cdot \dder{D}_2 \shift{(0,1)}
			\dder{D}_1' \cdot \dder{D}_0' \cdot \dder{D}_2 \shift{(1,2)}
			\dder{D}_1' \cdot \dder{D}_2' \cdot \dder{D}_0''$ then in the last derivation $\dder{D}_1'$ and $\dder{D}_2'$ are independent.
		\end{enumerate}
	\end{lemma}
\end{frame}

\begin{frame}{Results: the Three Steps Lemma}\justifying
Meaning of the previous the previous Lemma: \pause 
\begin{enumerate}
	\item if two derivation steps are independent and we are able to \emph{anticipate} something, then the steps remains independent; \pause 
	\item the independence of two derivation steps can depend on the steps before them, in particular \emph{posticipating} something does not preserves independence.
\end{enumerate}
\pause

\begin{block}{Remark}
The possibility of losing independence posticipating a derivation is due to left-linearity: it can be proved that in linear systems \emph{independence is global}, i.e. that given two independent contiguous steps, if  at the end of a sequence of switchings they are again contiguous, then they are also still independent (\cite{heindel2009category}).
\end{block}

\end{frame}

\begin{frame}{Risultati: Canonical Form}\justifying

	\begin{theorem}[No need of useless switches]
		Let $\der{D}$ and $\der{D}'$ be derivations If $\der{D} \shifteq[\sigma] \der{D}'$ then there is a switching sequence 
		%\begin{center}
		$\der{D} \shift{\nu_1} \der{D}_1 \shift{\nu_2} \ldots
		\shift{\nu_n} \der{D}'$
		%\end{center}
		consisting only of inversions for $\sigma$ (i.e. permutations $(i, i+1)$ such that $\sigma(i)>\sigma(i+1)$).
	\end{theorem}

\begin{theorem}[Canonical form]
	Let $\der{D}$ and $\der{D}'$ be derivations. If
	$\der{D} \shifteq[\sigma]\der{D}'$ then there is a switching sequence
	\begin{center}
		$\der{D} \shift{\nu_1} \der{D}_1 \shift{\nu_2} \ldots
		\shift{\nu_n} \der{D}'$
	\end{center}
	where for $i \in [1,n]$ we have $\nu_i  = \transp{k}$ with 
	$k = \max \{ j \mid \transp{j} \in \inv{\nu_{i,n}}\}$.
\end{theorem}
\end{frame}

\begin{frame}{Future works}\justifying 
	This work is only the first step in an attempt to construct a concurrent semantics for left-linear DPO rewriting systems.

	In \cite{baldan2017domains} a category of domains, called \emph{weak prime domains} has been proposed as a possible concurrent semantic for systems allowing fusions. We plan to build on the results proved in this paper to show that to every left-linear DPO rewriting system it is possible to associate one of these domains, representing ``derivation up to switches''.
\end{frame}

\begin{frame}{\hspace{1pt}}
	\LARGE
	\begin{center}
		Thank you for your attention!
	\end{center}

\end{frame}

\section{Bibliography}
\begin{frame}[allowframebreaks]{Bibliography}\justifying 
\printbibliography
\end{frame}

\end{document}