### 2024-09-09 concur/presentazione/Figure/coffee-rewriting2.tex

\begin{center}

  %%% REWRITING 
  % 
  %
  % 
  \begin{tikzpicture}[node distance=8mm, font=\small]
    % graph G0
    \node (G0) {
      \begin{tikzpicture}[node distance=2mm, label distance=1mm, font=\scriptsize]
        % 
        \node at (0,1.0) [node, label=left:$1$] (1) {}
        edge [in=110, out=140, loop] node[above] {$\mathit{m}$} ()
        edge [in=220, out=250, loop] node[below] {$\mathit{c}$} ();

        \node at (0.7,1.0) [node, label=right:$2$] (2) {}
        edge [in=40, out=70, loop] node[above] {$\mathit{u}$} ();

        \draw[->] (1) to node[above, pos=0.4]{$s$} (2);        
      %
      \pgfBox
      \end{tikzpicture} 
    };
    \node[below] at (G0.south) {$G_0$};
    % graph G1
    \node [right=of G0] (G1) {
      \begin{tikzpicture}[node distance=2mm, label distance=1mm,, font=\scriptsize]
      % 
      \node at (0.6,1.0) [node, label=right:$12$] (12) {}
        edge [in=220, out=250, loop] node[below] {$\mathit{c}$} ()
        edge [in=40, out=70, loop] node[above] {$\mathit{u}$} ()
        edge [in=300, out=330, loop] node[below] {$\mathit{s}$} ();        
      %
      \pgfBox
      \end{tikzpicture}
    };
    \node[below] at (G1.south) {$G_1$};

    % graph G2
    \node [right=of G1] (G2) {        
      \begin{tikzpicture}[node distance=2mm, label distance=1mm, font=\scriptsize]
        %
        \node at (0.4,0.5) {};
      \node at (0.6,1.0) [node, label=right:$12$] (12) {}
%        edge [in=220, out=250, loop] node[below] {$\mathit{b}$} ()
        edge [in=40, out=70, loop] node[above] {$\mathit{u}$} ()
        edge [in=300, out=330, loop] node[below] {$\mathit{s}$} ();        
      %
      \pgfBox
    \end{tikzpicture}
    };
    \node[below] at (G2.south) {$G_2$};

    % graph G3
    \node [right=of G2] (G3) {      
      \begin{tikzpicture}[node distance=2mm, label distance=1mm, font=\scriptsize]
        %
        \node at (0.4,0.5) {};
      \node at (0.6,1.0) [node, label=right:$12$] (12) {}
%        edge [in=220, out=250, loop] node[below] {$\mathit{b}$} ()
        edge [in=40, out=70, loop] node[above] {$\mathit{r}$} ()
        edge [in=300, out=330, loop] node[below] {$\mathit{s}$} ();        
        % 
        \pgfBox
      \end{tikzpicture}
    };
    \node[below] at (G3.south) {$G_3$};

    \path (G0) edge[->] node[trans, above] {$\rho_m$} (G1);
    \path (G1) edge[->] node[trans, above] {$\rho_c$} (G2);
    \path (G2) edge[->] node[trans, above] {$\rho_s$} (G3);    
  \end{tikzpicture}
\end{center}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../final"
%%% End: