### 2024-09-09 concur/presentazione/Figure/coffee-rules2.tex

\begin{center}

  %%% RULES 
  % 
  % RULE add white sugar
  % 
  \begin{tikzpicture}[node distance=3mm, font=\small]
    %, baseline=(current bounding box.center), font=\small]
      \node (l) {
      \begin{tikzpicture}[node distance=2mm, label distance=1mm, font=\scriptsize]
      % 
      \node at (0,1.0) [node, label=left:$1$] (1) {}
        edge [in=110, out=140, loop] node[above] {$\mathit{m}$} ();

        \node at (0,0.6) {};

        \node at (0.7,1.0) [node, label=right:$2$] (2) {};

        \draw[->] (1) to node[above, pos=0.4]{$\mathit{s}$} (2);        
      %
      \pgfBox
      \end{tikzpicture} 
    };
    \node [right=of l] (r) {
      \begin{tikzpicture}[node distance=2mm, label distance=1mm, font=\scriptsize]
        %
        \node at (0,1.45) {};
        \node at (0,0.5) {};
        \node at (0.6,1.0) [node, label=right:$12$] (12) {}
          edge [in=300, out=330, loop] node[below] {$\mathit{s}$} (); 
      %
      \pgfBox
      \end{tikzpicture}
    };
    %\node[below] at (l.south east) {$\rho_w$};
    \path (l) edge[->] (r);
    \node at ($(l)!0.54!(r)$) [below=7.5mm] {$\rho_m$: add milk and sugar}; 
    \end{tikzpicture}
    % 
    %
    \hspace{5mm}
    % \hfill
    % 
    % RULE cocoa
    %
    \begin{tikzpicture}[node distance=3mm, font=\small]
    %, baseline=(current bounding box.center), font=\small]
      \node (l) {
      \begin{tikzpicture}[node distance=2mm, label distance=1mm, font=\scriptsize]
      % 
       \node at (0,1.0) [node, label=left:$1$] (1) {}
        edge [in=220, out=250, loop] node[below] {$\mathit{c}$} ();

        \node at (0,0.6) {};

        \node at (0.7,1.0) [node, label=right:$2$] (2) {};

        \draw[->] (1) to node[above, pos=0.4]{$\mathit{s}$} (2);

      %
      \pgfBox
      \end{tikzpicture} 
    };
    \node [right=of l] (r) {
      \begin{tikzpicture}[node distance=2mm, label distance=1mm, font=\scriptsize]
      %
      \node at (0,1.45) {};  
      \node at (0,0.5) {};
      \node at (0.6,1.0) [node, label=right:$12$] (12) {}
          edge [in=300, out=330, loop] node[below] {$\mathit{s}$} ();        
      %
      \pgfBox
      \end{tikzpicture}
    };
    \path (l) edge[->]  (r);
    \node at ($(l)!0.54!(r)$) [below=7.5mm] {$\rho_c$: add cocoa and sugar};         
    \end{tikzpicture}
    %
    \hspace{5mm}
    % RULE ready
    % 
    \begin{tikzpicture}[node distance=3mm, font=\small]
    %, baseline=(current bounding box.center), font=\small]
      \node (l) {
      \begin{tikzpicture}[node distance=2mm, label distance=1mm, font=\scriptsize]
      % 
      \node at (0.4,0.5) {};
      \node at (0.6,1.0) [node, label=right:$1$] (1) {}
        edge [in=40, out=70, loop] node[above] {$\mathit{u}$} ()
        edge [in=300, out=330, loop] node[below] {$\mathit{s}$} ();
      %
      \pgfBox
      \end{tikzpicture} 
    };
    \node [right=of l] (r) {
      \begin{tikzpicture}[node distance=2mm, label distance=1mm, font=\scriptsize]
      % 
      \node at (0.4,0.5) {};
      \node at (0.6,1.0) [node, label=right:$1$] (1) {}
        edge [in=40, out=70, loop] node[above] {$\mathit{r}$} ()
        edge [in=300, out=330, loop] node[below] {$\mathit{s}$} ();
      %
      \pgfBox
      \end{tikzpicture}
    };
    \path (l) edge[->] (r);    
    \node at ($(l)!0.54!(r)$) [below=7.5mm] {$\rho_s$: stir and drink!}; 
    \end{tikzpicture}
  %
\end{center}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../final"
%%% End: