### 2025-02-28 2025-notes.tex
notes

\documentclass{article}
\usepackage{pdfpages}
\usepackage{hyperref}
\usepackage{forloop}
\usepackage{tikz}

\newcounter{pdfpage}
\begin{document}

\hypertarget{start}{}

\section*{Some ideas for the second part of the paper}

\begin{itemize}

  \thispagestyle{empty}

\item Irreducibles and axioms:
  \begin{itemize}
  \item Characterisation of the irreducibles \hyperlink{pdfpage.1}{page 1}
  \item Interchangebility axioms in this context \hyperlink{pdfpage.3}{page 3}
  \end{itemize}

\item \textbf{Idea}: work with rewriting systems with stronger
  properties ensuring both interchangeability and good properties
  w.r.t. fusions (\hyperlink{pdfpage.4}{page 4}, done also in the paper)

  \emph{Paradigmantic example}: left-linear rules on graphs where lhs
  has no isolated nodes and right-leg is inective on edges.

  The context ensures that 
  \begin{itemize}
  \item the rewriting system is well swiitching

  \item morphisms into the colimit are in the class of morphisms used for lhs

  \item existence of a minimal quotient to which a rule applies
  \end{itemize}

  Additionally, in this setting:
  \begin{itemize}
  \item a rewriting step can be lifted up and down along quotients (see property 1 page \hyperlink{pdfpage.9}{page 9}, property 2 page \hyperlink{pdfpage.10}{page 10})

  \item independence (up to fusion) can be checked in the colimit
    [\hyperlink{pdfpage.13}{page 13}, some considerations here might be uselessly
    complex as they might be done before taking the assumptions on
    rules]

  \item 
   we can work with derivations up to fusion: each step can be preceded by a quotient which is (the minimal one - unclear whether it is convenient to require minimality) sufficient to allow the application of the rule [some reflections starting from page \hyperlink{pdfpage.17}{page 17}, including a notion of switch and further considerations on independence in the colimit)

 \item characterise transitively interchangeable irreducibles via
   derivations up to fusions [\hyperlink{pdfpage.23}{page 23}]

 \item prove the interchangeability axioms [page
   \hyperlink{pdfpage.24}{page 24}]

 \end{itemize}
\end{itemize}

\setcounter{page}{0}

\forloop{pdfpage}{1}{\value{pdfpage} < 27}{
  \includepdf[pages=\thepdfpage,pagecommand={
    \vspace*{-9cm}
    \hypertarget{pdfpage.\thepdfpage}{}
        \begin{tikzpicture}[remember picture, overlay]
            % Position text at the top-right corner
            \node[anchor=north east, xshift=-10pt, yshift=-10pt] at (current page.north east) {
                \textbf{\large \hyperlink{start}{Start}}
            };
        \end{tikzpicture}
        }]{2025-notes-file.pdf}
}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: